---
title: "preparing data"
author: "anne"
date: "13 desember 2018"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## ---------------------- Plots/ environment-------------------------------

#### Look at data structure
```{r}
head(data_plots, n = 3)

str(data_plots)


```


#### Deal with NAs (! not yet, we will do it later since we want to add new columns with bind_col and then the two dataframes need the same length !)
```{r}
# nrow(data_plots)
# nrow(na.omit(data_plots))
# 
# unique(unlist(lapply(data_plots, function (x) which(is.na(x)))))
# data_plots[102, ] # true NAs (fieldlayer hight, shrublayer height 1 and 2 missing)
# data_plots[108, ] # true NAs (fieldlayer hight missing)
# data_plots[217, ] # true NAs (fieldlayer hight missing)
# data_plots[90, ]  # true NAs (no canopy cover measures)
# data_plots[275, ] # true NAs (missing grain size)
# data_plots[276, ] # true NAs (missing grain size)
# data_plots[279, ] # true NAs (missing grain size)
# data_plots[185, ] # true NAs (missing soil dry weight and loi)
# data_plots[186, ] # true NAs (missing soil dry weight and loi)
# data_plots[35, ] # true NAs (missing loi)
# 
# data_plots <- na.omit(data_plots)
# nrow(data_plots)
# nrow(na.omit(data_plots))

```


#### change some character variables to factor
```{r}
data_plots$site <- as.factor(data_plots$site)
str(data_plots$site)

data_plots$method <- as.factor(data_plots$method)
str(data_plots$method)

data_plots$exposure <- as.factor(data_plots$exposure)
str(data_plots$exposure)

data_plots$grain_size_stand_f <- as.factor(data_plots$grain_size_stand)
str(data_plots$grain_size_stand_f)

data_plots$grain_size_f <- as.factor(data_plots$grain_size)
str(data_plots$grain_size_f)

```


#### Change level names in exposure 
```{r}
levels(data_plots$exposure)[levels(data_plots$exposure)=="0"] <- "F"
levels(data_plots$exposure)[levels(data_plots$exposure)=="S/E"] <- "SE"
levels(data_plots$exposure)

```


#### Add a column with only year
```{r}
library(tidyverse)

data_plots <- data_plots %>%
  separate(date, into = c("day", "month", "year"), remove = F) # Can also use  sep = "."  to specify the seperator

data_plots$year <- as.numeric(data_plots$year)
data_plots$month <- as.numeric(data_plots$month)
data_plots$day <- as.numeric(data_plots$day)

```


#### Add a column with site ID and one with plot number
```{r}
library(tidyverse)

data_plots <- data_plots %>%
  separate(ID, into = c("siteID", "plotnumb"), remove = F) # Can also use  sep = "."  to specify the seperator

data_plots$siteID <- as.factor(data_plots$siteID)
data_plots$plotnumb <- as.factor(data_plots$plotnumb)
  
  
```


#### Add a column with reference plot yes or no
```{r}

data_plots$plotnumb <- as.numeric(data_plots$plotnumb)
data_plots$reference <- ifelse(data_plots$plotnumb >= 9, "yes", "no")

data_plots$reference <- as.factor(data_plots$reference)


```


#### Rename variable site
```{r}
data_plots <- data_plots %>% 
  rename(
    site_type = site)

```


#### Add mean and sd species cover to plots dataset
```{r}
load("processdata/data_plots_2020_04_20.rda")

data_plots2 <- bind_cols(data_plots, data_species_wide2$m)
data_plots2 <- data_plots2 %>% 
  rename(
    mean_cover = ...44)

data_plots2 <- bind_cols(data_plots2, data_species_wide2$sd)
data_plots2 <- data_plots2 %>% 
  rename(
    sd_cover = ...45)

```

#### Deal with NAs (after adding new columns)
```{r}
nrow(data_plots2)
nrow(na.omit(data_plots2))

unique(unlist(lapply(data_plots2, function (x) which(is.na(x)))))
data_plots2[102, ] # true NAs (fieldlayer hight, shrublayer height 1 and 2 missing)
data_plots2[108, ] # true NAs (fieldlayer hight missing)
data_plots2[217, ] # true NAs (fieldlayer hight missing)
data_plots2[90, ]  # true NAs (no canopy cover measures)
data_plots2[275, ] # true NAs (missing grain size)
data_plots2[276, ] # true NAs (missing grain size)
data_plots2[279, ] # true NAs (missing grain size)
data_plots2[185, ] # true NAs (missing soil dry weight and loi)
data_plots2[186, ] # true NAs (missing soil dry weight and loi)
data_plots2[35, ] # true NAs (missing loi)

data_plots2 <- na.omit(data_plots2)
nrow(data_plots2)
nrow(na.omit(data_plots2))

```


#### Add a total cover column from adding up the cover of each layer
```{r}
data_plots2$total_cover_layers <- data_plots2$ground_layer_c + data_plots2$field_layer_c + data_plots2$shrub_layer_c + data_plots2$tree_layer_c
```


#### Subset all reference plots
```{r}
data_plots2_reference <- filter(data_plots2, reference == "yes")
data_plots2_reference <- as.data.frame(data_plots2_reference)

```


#### Subset all non-reference plots
```{r}
data_plots2_normal <- filter(data_plots2, reference == "no")
data_plots2_normal <- as.data.frame(data_plots2_normal)

```


#### Save changes of data
```{r}
save(data_plots, file = "data_plots_2020_06_25.rda")
save(data_plots2, file = "data_plots2_2020_06_29.rda") # Like data_plots, but inluding column with mean cover and sd (species), total_cover_layers
save(data_plots2_reference, file = "data_plots2_reference_2020_06_29.rda") # Like data_plots2, but only reference plots
save(data_plots2_normal, file = "data_plots2_normal_2020_06_29.rda") # Like data_plots2, but only non-reference plots

load("processdata/data_plots_2020_06_25.rda")
load("processdata/data_plots2_2020_06_29.rda") # Like data_plots, but inluding column with mean cover and sd (species)
load("processdata/data_plots2_reference_2020_06_29.rda") # Like data_plots2, but only reference plots
load("processdata/data_plots2_normal_2020_06_29.rda") # Like data_plots2, but only non-reference plots
```








## ---------------------- Species/ cover ---------------------------

#### Look at data structure
```{r}
head(data_species, n = 3)

str(data_species)


```


#### Deal with NAs
```{r}
nrow(data_species)
nrow(na.omit(data_species)) # 0 NAs

# unique(unlist(lapply(data_species, function (x) which(is.na(x)))))
# data_species <- na.omit(data_species)
# nrow(data_species)
# nrow(na.omit(data_species))


nrow(data_species_wide)
nrow(na.omit(data_species_wide)) # 0 NAs


```


#### Spread the species data into wide format (species as columns)
```{r}
# library(tidyr)
# 
# data_species_wide <- data_species %>% pivot_wider(names_from = species, values_from = cover, names_repair = "minimal", values_fill = list(cover = 0), values_fn = list(cover = sum))
# 
# data_species_wide <- as.data.frame(data_species_wide)


```


#### Add a column with site ID and one with plot number
```{r}
library(tidyverse)

data_species_wide <- data_species_wide %>%
  separate(ID, into = c(NA, "plotID"), remove = F) # Can also use  sep = "."  to specify the seperator , convert = TRUE
data_species_wide <- data_species_wide %>%
  separate(ID, into = c("siteID", NA), remove = F) # Can also use  sep = "."  to specify the seperator , convert = TRUE

data_species_wide$siteID <- as.factor(data_species_wide$siteID)

  
  
```


#### Add a column with reference plot yes or no
```{r}
data_species_wide$plotID <- as.numeric(data_species_wide$plotID)
data_species_wide$reference <- ifelse(data_species_wide$plotID > 8, "yes", "no")

data_species_wide$reference <- as.factor(data_species_wide$reference)
data_species_wide$plotID <- as.factor(data_species_wide$plotID)

str(data_species_wide$plotID)
str(data_species_wide$siteID)
str(data_species_wide$reference)

head(data_species_wide$plotID, n = 25L)
head(data_species_wide$reference, n = 25L)

```

#### Remove space and dots in species names
```{r}
library(janitor)

data_species_wide <- data_species_wide %>%
  clean_names()

```



#### Summarise over columns
```{r}
## create a list with function to use over each column
mean_sd <- list(
  mean = ~mean(.x, na.rm = TRUE), 
  sd = ~sd(.x, na.rm = TRUE)
)

## Creat new dataframe with summary for each species column (mean and sd)

#### per site
data_species_wide_site_mean_sd <- data_species_wide %>%
  group_by(site_id) %>%
  summarise(
    across(anemone_nemorosa:lolium_perenne, mean_sd),
     
  )


#### over entire df
data_species_wide_mean <- data_species_wide %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, mean),
     
  )

data_species_wide_sd <- data_species_wide %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, sd),
     
  )

## If preferred, add the rows to (existing) dataframe
data_species_wide_temp <- bind_rows(data_species_wide, data_species_wide_mean)
data_species_wide_2 <- bind_rows(data_species_wide_temp, data_species_wide_sd)


```

#### Summarise over rows
```{r}
# use c_across() to more easily select many variables
data_species_wide2 <- data_species_wide2 %>% rowwise() %>% 
  mutate(m = mean(c_across(taraxacum_sp:lolium_perenne))) %>%
  mutate(sd = sd(c_across(taraxacum_sp:lolium_perenne)))

```

#### Subset all reference plots
```{r}
data_species_wide2_reference <- filter(data_species_wide2, reference == "yes")
data_species_wide2_reference <- as.data.frame(data_species_wide2_reference)

```


#### Summarise over columns in reference data
```{r}
data_species_wide2_reference_mean <- data_species_wide2_reference %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, mean),
     
  )

##### Why the hell does this not work??

# data_species_wide2_reference_sd <- data_species_wide2_reference %>%
#   summarise(
#     across(taraxacum_sp:lolium_perenne, sd),
#      
#   ) 


data_species_wide2_reference <- bind_rows(data_species_wide2_reference, data_species_wide2_reference_mean)
# data_species_wide2_reference <- bind_rows(data_species_wide2_reference, data_species_wide2_reference_sd)

```

#### Subset all non-reference plots
```{r}
data_species_wide2_normal <- filter(data_species_wide2, reference == "no")
data_species_wide2_normal <- as.data.frame(data_species_wide2_normal)

```


#### Summarise over columns in reference data
```{r}
data_species_wide2_normal_mean <- data_species_wide2_normal %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, mean),
     
  )

##### Why the hell does this not work??

# data_species_wide2_normal_sd <- data_species_wide2_normal %>%
#   summarise(
#     across(taraxacum_sp:lolium_perenne, sd),
# 
#   )


data_species_wide2_normal <- bind_rows(data_species_wide2_normal, data_species_wide2_normal_mean)
# data_species_wide2_reference <- bind_rows(data_species_wide2_reference, data_species_wide2_reference_sd)

```


#### ID as factor (for reference dataset)
```{r}
data_species_wide2_reference$id <- as.factor(data_species_wide2_reference$id)

```



#### Save new data/ changes
```{r}
save(data_species_wide, file = "data_species_wide_2020_06_26.rda")
save(data_species_wide2, file = "data_species_wide2_2020_06_26.rda") # Includes mean and sd of species for both columns and rows
save(data_species_wide2_reference, file = "data_species_wide2_reference_2020_06_29.rda") # Only reference plots, includes mean adn sd for rows and mean for columns
save(data_species_wide2_normal, file = "data_species_wide2_normal_2020_06_29.rda") # Only non-reference plots, includes mean adn sd for rows and mean for columns

load("processdata/data_species_wide_2020_06_26.rda")
load("processdata/data_species_wide2_2020_06_26.rda") # Includes mean and sd of species for columns and rows
load("processdata/data_species_wide2_reference_2020_06_29.rda") # Only reference plots, includes mean adn sd for rows and mean for columns
load("processdata/data_species_wide2_normal_2020_06_29.rda") # Only non-reference plots, includes mean adn sd for rows and mean for columns

head(data_species_wide)

```



















