---
title: "preparing data"
author: "anne"
date: "13 desember 2018"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## ---------------------- Plots/ environment-------------------------------

#### Look at data structure
```{r}
head(data_plots, n = 3)

str(data_plots)


```


#### Deal with NAs (! not yet, we will do it later since we want to add new columns with bind_col and then the two dataframes need the same length !)
```{r}
# nrow(data_plots)
# nrow(na.omit(data_plots))
# 
# unique(unlist(lapply(data_plots, function (x) which(is.na(x)))))
# data_plots[102, ] # true NAs (fieldlayer hight, shrublayer height 1 and 2 missing)
# data_plots[108, ] # true NAs (fieldlayer hight missing)
# data_plots[217, ] # true NAs (fieldlayer hight missing)
# data_plots[90, ]  # true NAs (no canopy cover measures)
# data_plots[275, ] # true NAs (missing grain size)
# data_plots[276, ] # true NAs (missing grain size)
# data_plots[279, ] # true NAs (missing grain size)
# data_plots[185, ] # true NAs (missing soil dry weight and loi)
# data_plots[186, ] # true NAs (missing soil dry weight and loi)
# data_plots[35, ] # true NAs (missing loi)
# 
# data_plots <- na.omit(data_plots)
# nrow(data_plots)
# nrow(na.omit(data_plots))

```


#### change some character variables to factor
```{r}
#### data set "data_plots" ----
data_plots$site <- as.factor(data_plots$site)
str(data_plots$site)

data_plots$method <- as.factor(data_plots$method)
str(data_plots$method)

data_plots$exposure <- as.factor(data_plots$exposure)
str(data_plots$exposure)

data_plots$grain_size_stand_f <- as.factor(data_plots$grain_size_stand)
str(data_plots$grain_size_stand_f)

data_plots$grain_size_f <- as.factor(data_plots$grain_size)
str(data_plots$grain_size_f)


#### data set "data_plots2_normal" ----
# data_plots2_normal$plotnumb <- as.factor(data_plots2_normal$plotnum)
# str(data_plots2_normal$plotnum)

```


#### Change level names in exposure 
```{r}
levels(data_plots$exposure)[levels(data_plots$exposure)=="0"] <- "F"
levels(data_plots$exposure)[levels(data_plots$exposure)=="S/E"] <- "SE"
levels(data_plots$exposure)

```



#### Add a column with site ID and one with plot number
```{r}
library(tidyverse)

data_plots <- data_plots %>%
  separate(ID, into = c("site_id", "plot_id"), remove = F) # Can also use  sep = "."  to specify the seperator

data_plots$site_id <- as.factor(data_plots$site_id)

  
  
```


#### Add a column with reference plot yes or no
```{r}

# data_plots$plot_id <- as.numeric(data_plots$plot_id)
data_plots$reference <- ifelse(data_plots$plot_id >= 9, "yes", "no")

data_plots$reference <- as.factor(data_plots$reference)
str(data_plots$reference)

data_plots$plot_id <- as.factor(data_plots$plot_id)

```


#### Rename variables
```{r}
data_plots <- data_plots %>% 
  rename(
    site_type = site)

data_plots <- data_plots %>% 
  rename(
    id = ID)

```


#### Add mean and sd species cover to plots dataset
```{r}
species_cover_meansd <- data_species_wide2[, c(1, 194:195)]

data_plots2 <- data_plots %>% 
  left_join(species_cover_meansd) %>%
  as.data.frame()

```

#### Deal with NAs (after adding new columns)
```{r}
nrow(data_plots2)
nrow(na.omit(data_plots2))

unique(unlist(lapply(data_plots2, function (x) which(is.na(x)))))
data_plots2[102, ] # true NAs (fieldlayer hight, shrublayer height 1 and 2 missing)
data_plots2[108, ] # true NAs (fieldlayer hight missing)
data_plots2[217, ] # true NAs (fieldlayer hight missing)
data_plots2[90, ]  # true NAs (no canopy cover measures)
data_plots2[275, ] # true NAs (missing grain size)
data_plots2[276, ] # true NAs (missing grain size)
data_plots2[279, ] # true NAs (missing grain size)
data_plots2[185, ] # true NAs (missing soil dry weight and loi)
data_plots2[186, ] # true NAs (missing soil dry weight and loi)
data_plots2[35, ] # true NAs (missing loi)
data_plots2[198, ] # No species cover mean and sd, so only important for that really

data_plots2 <- na.omit(data_plots2)
nrow(data_plots2)
nrow(na.omit(data_plots2))

```


#### Add a total cover column from adding up the cover of each layer
```{r}
data_plots2$total_cover_layers <- data_plots2$ground_layer_c + data_plots2$field_layer_c + data_plots2$shrub_layer_c + data_plots2$tree_layer_c
```


#### Subset all reference plots
```{r}
data_plots2_reference <- filter(data_plots2, reference == "yes")
data_plots2_reference <- as.data.frame(data_plots2_reference)

```


#### Subset all non-reference plots
```{r}
data_plots2_normal <- filter(data_plots2, reference == "no")
data_plots2_normal <- as.data.frame(data_plots2_normal)

```



#### Rename variables
```{r}
data_plots2 <- data_plots2 %>% 
  rename(
    mean_cover = m)

data_plots2 <- data_plots2 %>% 
  rename(
    sd_cover = sd)

```



#### Add columns with species richness and diversity from the species dataframes to the plots dataframes
Start with the dataframe which includes both normal and reference plots
```{r}

species_diversity <- data_species_wide3[, c(1, 196:199)]

data_plots3 <- data_plots2 %>% 
  left_join(species_diversity) %>%
  as.data.frame()

sum(is.na(data_plots3))

```



#### Order factor levels, so that the most used one is first
```{r}

plot(data_plots3$site_type)
data_plots3$site_type <- factor(data_plots3$site_type, levels = c("road", "crossing"))

plot(data_plots3$grain_size_stand_f)
data_plots3$grain_size_stand_f <- factor(data_plots3$grain_size_stand_f, levels = c("6", "1", "2", "3", "5"))

```





#### Save changes of data
```{r}
save(data_plots, file = "data_plots_2020_06_25.rda")
save(data_plots2, file = "data_plots2_2020_06_29.rda") # Like data_plots, but inluding column with mean cover and sd (species), total_cover_layers
save(data_plots2_reference, file = "data_plots2_reference_2020_07_16.rda") # Like data_plots2, but only reference plots
save(data_plots2_normal, file = "data_plots2_normal_2020_07_16.rda") # Like data_plots2, but only normal plots

save(data_plots, file = "data_plots_2020_07_17.rda")
save(data_plots2, file = "data_plots2_2020_07_17.rda") # Like data_plots, but inluding column with mean cover and sd (species), total_cover_layers
save(data_plots2_reference, file = "data_plots2_reference_2020_07_17.rda") # Like data_plots2, but only reference plots
save(data_plots2_normal, file = "data_plots2_normal_2020_07_17.rda") # Like data_plots2, but only normal plots

save(data_plots3, file = "data_plots3_2020_07_16.rda") # Like data_plots2, but species richness and diversity
save(data_plots3, file = "data_plots3_2020_07_17.rda") # Like data_plots2, but species richness and diversity


load("processdata/data_plots_2020_06_25.rda")
load("processdata/data_plots2_2020_07_17.rda") # Like data_plots, but inluding column with mean cover and sd (species cover)
load("processdata/data_plots2_reference_2020_07_17.rda") # Like data_plots2, but only reference plots
load("processdata/data_plots2_normal_2020_07_17.rda") # Like data_plots2, but only normal (non-reference) plots
load("processdata/data_plots3_2020_07_16.rda") # Like data_plots2, but species richness and diversity
load("processdata/data_plots3_2020_07_17.rda") # Like data_plots2, but species richness and diversity

```








## ---------------------- Species/ cover ---------------------------

#### Look at data structure
```{r}
head(data_species, n = 3)

str(data_species)


```


#### Deal with NAs
```{r}
nrow(data_species)
nrow(na.omit(data_species)) # 0 NAs

# unique(unlist(lapply(data_species, function (x) which(is.na(x)))))
# data_species <- na.omit(data_species)
# nrow(data_species)
# nrow(na.omit(data_species))


nrow(data_species_wide)
nrow(na.omit(data_species_wide)) # 0 NAs


```


#### Spread the species data into wide format (species as columns)
```{r}
# library(tidyr)
# 
# data_species_wide <- data_species %>% pivot_wider(names_from = species, values_from = cover, names_repair = "minimal", values_fill = list(cover = 0), values_fn = list(cover = sum))
# 
# data_species_wide <- as.data.frame(data_species_wide)


```


#### Add a column with site ID and one with plot number
```{r}
library(tidyverse)

data_species_wide <- data_species_wide %>%
  separate(id, into = c(NA, "plot_id"), remove = F) # Can also use  sep = "."  to specify the seperator , convert = TRUE
data_species_wide <- data_species_wide %>%
  separate(id, into = c("site_id", NA), remove = F) # Can also use  sep = "."  to specify the seperator , convert = TRUE

data_species_wide$plot_id <- as.factor(data_species_wide$plot_id)
data_species_wide$site_id <- as.factor(data_species_wide$site_id)

  
  
```


#### Add a column with reference plot yes or no
```{r}
data_species_wide$plot_id <- as.numeric(data_species_wide$plot_id)
data_species_wide$reference <- ifelse(data_species_wide$plot_id > 8, "yes", "no")

data_species_wide$reference <- as.factor(data_species_wide$reference)
data_species_wide$plot_id <- as.factor(data_species_wide$plot_id)

str(data_species_wide$plot_id)
str(data_species_wide$site_id)
str(data_species_wide$reference)

head(data_species_wide$plot_id, n = 25L)
head(data_species_wide$reference, n = 25L)

```

#### Remove spaces, dots and capital letters in variable names
```{r}
# library(janitor)
# 
# data_species_wide <- data_species_wide %>%
#   clean_names()

```



#### Summarise cover for each column
```{r}
## create a list with function to use over each column
mean_sd <- list(
  mean = ~mean(.x, na.rm = TRUE), 
  sd = ~sd(.x, na.rm = TRUE)
)

## Creat new dataframe with summary for each species column (mean and sd)

#### per site
data_species_wide_site_mean_sd <- data_species_wide %>%
  group_by(site_id) %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, mean_sd),
     
  )


#### over entire df
data_species_wide_mean <- data_species_wide %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, mean),
     
  )

data_species_wide_sd <- data_species_wide %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, sd),
     
  )

#### If preferred, add the rows to (existing) dataframe:

# data_species_wide_temp <- bind_rows(data_species_wide, data_species_wide_mean)
# data_species_wide_2 <- bind_rows(data_species_wide_temp, data_species_wide_sd)


```

#### Summarise cover for each row (plot/ site)
```{r}
# use c_across() to more easily select many variables
data_species_wide2 <- data_species_wide %>% rowwise() %>% 
  mutate(mean_cover = mean(c_across(taraxacum_sp:lolium_perenne))) %>%
  mutate(sd_cover = sd(c_across(taraxacum_sp:lolium_perenne)))

```






#### Rename variables
```{r}
data_species_wide3 <- data_species_wide3 %>% 
  rename(
    mean_plot_cover = m)

data_species_wide3 <- data_species_wide3 %>% 
  rename(
    sd_plot_cover = sd)

```






#### Subset all reference plots
```{r}
#### Use latest dataframe here ----
data_species_wide3_reference <- filter(data_species_wide3, reference == "yes")
data_species_wide3_reference <- as.data.frame(data_species_wide3_reference)


data_species_wide3_sites_reference <- filter(data_species_wide3_sites, reference == "yes")
data_species_wide3_sites_reference <- as.data.frame(data_species_wide3_sites_reference)

```


#### Subset all restored plots
```{r}
#### Use latest dataframe here ----
data_species_wide3_normal <- filter(data_species_wide3, reference == "no")
data_species_wide3_normal <- as.data.frame(data_species_wide3_normal)

```


#### Save new data/ changes
```{r}
save(data_species_wide, file = "data_species_wide_2020_06_26.rda")

# Includes columns with species cover mean and sd:
# save(data_species_wide2, file = "data_species_wide2_2020_06_26.rda") 
save(data_species_wide2, file = "data_species_wide2_2020_07_15.rda")


# Only reference plots - Includes columns with mean and sd for species cover, species richness measures and species diversity measures:
save(data_species_wide3_reference, file = "data_species_wide3_reference_2020_07_15.rda")
# Only normal (non-reference) plots - Includes columns with mean and sd for species cover, species richness measures and species diversity measures:
save(data_species_wide3_normal, file = "data_species_wide3_normal_2020_07_15.rda")



load("processdata/data_species_wide_2020_06_26.rda")

#### OBS! somehow the plot and site id columns are wrong in this version...
# # Includes columns with species cover mean and sd:
# load("processdata/data_species_wide2_2020_07_15.rda") 

# Like data_species_wide2 but grouped by site id:
load("processdata/data_species_wide2_sites_2020_07_15.rda")


head(data_species_wide)

```






## ---------------------- Full data ---------------------------

#### Bring plot and species datasets to same length
```{r}
library(dplyr)

data_full <- inner_join(data_plots4, data_species_wide4, by = "id")


save(data_full, file = "data_full_2020_08_19.rda")


```

#### Make new colomn for region
```{r}
data_full$region <- ifelse(data_full$site_id_new.x < 11, "N", NA)
data_full$region <- ifelse(data_full$site_id_new.x == 11, "N", data_full$region)
data_full$region <- ifelse(data_full$site_id_new.x == 12, "E", data_full$region)
data_full$region <- ifelse(data_full$site_id_new.x == 13, "E", data_full$region)
data_full$region <- ifelse(data_full$site_id_new.x == 14, "N", data_full$region)
data_full$region <- ifelse(data_full$site_id_new.x > 14, "W", data_full$region)

data_full$region <- as.factor(data_full$region)
levels(data_full$region)

save(data_full, file = "data_full_2020_08_24.rda")

```






## Make species matrix for ordination --> no plots with less than 5 species are included, not species with less than 5 occurences 
```{r}
species <- data_full[, c(56:244)]
str(species)
sum(rowSums(species) == 0)
sum(colSums(species) == 0)
which(colSums(species) == 0)
species <- species[, -c(123)]


species_richness <- rep(0, times=nrow(species))
for(i in 1:nrow(species)){
  species_richness[i] <- sum(species[i,] > 0)
}
species_richness <- as.data.frame(species_richness)
species_2 <- cbind(species, species_richness)

species_sub5 <- subset(species_2, species_richness >5)
sum(rowSums(species_sub5) == 0)
sum(colSums(species_sub5) == 0)
which(colSums(species_sub5) == 0)

species_sub5 <- species_sub5[, -c(61, 110, 119, 160, 186, 189)]
which(colSums(species_sub5) == 0)
names(species_sub5)


save(species_sub5, file = "species_sub5_2020_208_27.rda")




# To check if we get other results when keeping most of the observations (we exclude a lot of reference plots when doing species_richness >5)
species <- data_full[, c(56:244)]
str(species)
sum(rowSums(species) == 0)
sum(colSums(species) == 0)
which(colSums(species) == 0)
species <- species[, -c(123)]


species_richness <- rep(0, times=nrow(species))
for(i in 1:nrow(species)){
  species_richness[i] <- sum(species[i,] > 0)
}
species_richness <- as.data.frame(species_richness)
species_2 <- cbind(species, species_richness)

species_sub1 <- subset(species_2, species_richness >1)
sum(rowSums(species_sub1) == 0)
sum(colSums(species_sub1) == 0)



save(species_sub1, file = "species_sub1_2020_09_11.rda")


```




## Make env.var_sub5 dataset for ordination --> adjust environmental data (delete rows that were sorted out in the species dataset)
```{r}
env.var <- data_full[, c(6:43, 252)]
str(env.var)
env.var <- env.var[, c(1:3, 8, 26, 29:30, 35:36, 39)]
env.var$site_type <- as.factor(env.var$site_type)
env.var$method <- as.factor(env.var$method)
env.var$exposure <- as.factor(env.var$exposure)
env.var$year <- as.factor(env.var$year)
env.var$grain_size_stand_f <- as.factor(env.var$grain_size_stand_f)

env.var_2 <- cbind(env.var, species_richness)
env.var_sub5 <- subset(env.var_2, species_richness >1)
env.var_sub5 <- env.var_sub5[, -c(11)]
names(env.var_sub5)

save(env.var_sub5, file = "env.var_sub5_2020_208_27.rda")


str(env.var_sub5$grain_size_stand_f)
table(env.var_sub5$grain_size_stand_f)
 #  1   2   3   5   6
 # 27   0  13   2 177
env.var_sub5$grain_size_stand_f <- as.numeric(env.var_sub5$grain_size_stand_f)
env.var_sub5$grain_size_stand_f <- ifelse(env.var_sub5$grain_size_stand_f >= 4, 6, env.var_sub5$grain_size_stand_f)
table(env.var_sub5$grain_size_stand_f)
 #  1   3   6
 # 27  13 179
env.var_sub5$grain_size_stand_f <- as.factor(env.var_sub5$grain_size_stand_f)
str(env.var_sub5$grain_size_stand_f)
table(env.var_sub5$grain_size_stand_f)


str(env.var_sub5$exposure)
table(env.var_sub5$exposure)
  # E   F   N  NE  NW   S  SE  SW   W 
  # 9 124  31   1   6  19  13   6  10
levels(env.var_sub5$exposure)[levels(env.var_sub5$exposure)=="NE"] <- "N"
str(env.var_sub5)
table(env.var_sub5$exposure)


save(env.var_sub5, file = "env.var_sub5_2020_09_08.rda")







# To check if we get other results when keeping most of the observations (we exclude a lot of reference plots when doing species_richness >5)
env.var_2 <- cbind(env.var, species_richness)
env.var_sub1 <- subset(env.var_2, species_richness >1)
env.var_sub1 <- env.var_sub1[, -c(11)]
str(env.var_sub1)
names(env.var_sub1)

save(env.var_sub1, file = "env.var_sub1_2020_09_11.rda")


str(env.var_sub1$grain_size_stand_f)
table(env.var_sub1$grain_size_stand_f)
 #  1   2   3   5   6 
 # 33   1  16   2 230 
env.var_sub1$grain_size_stand_f <- as.numeric(env.var_sub1$grain_size_stand_f)
env.var_sub1$grain_size_stand_f <- ifelse(env.var_sub1$grain_size_stand_f >= 4, 6, env.var_sub1$grain_size_stand_f)
env.var_sub1$grain_size_stand_f <- ifelse(env.var_sub1$grain_size_stand_f == 2, 1, env.var_sub1$grain_size_stand_f)
table(env.var_sub1$grain_size_stand_f)
 #  1   3   6 
 # 34  16 232 
env.var_sub1$grain_size_stand_f <- as.factor(env.var_sub1$grain_size_stand_f)
str(env.var_sub1$grain_size_stand_f)
table(env.var_sub1$grain_size_stand_f)


str(env.var_sub1$exposure)
table(env.var_sub1$exposure)
 #  E   F   N  NE  NW   S  SE  SW   W 
 # 13 166  33   6   7  23  17   6  11


save(env.var_sub1, file = "env.var_sub1_2020_09_11.rda")
```

#### Change levels for method, so that reference is in intercept
```{r}
levels(env.var_sub5$method)
env.var_sub5$method <- factor(env.var_sub5$method, levels = c("ref", "nat",  "pn",  "seed"))
plot(env.var_sub5$method)

# To check if we get other results when keeping most of the observations (we exclude a lot of reference plots when doing species_richness >5)
levels(env.var_sub1$method)
env.var_sub1$method <- factor(env.var_sub1$method, levels = c("ref", "nat",  "pn",  "seed"))
plot(env.var_sub1$method)

```


#### Make a variable years since restoration
```{r}
levels(env.var_sub5$year)

env.var_sub5$years_since <- ifelse(env.var_sub5$year == "2003", "17", NA)
env.var_sub5$years_since <- ifelse(env.var_sub5$year == "2007", "13", env.var_sub5$years_since)
env.var_sub5$years_since <- ifelse(env.var_sub5$year == "2009", "11", env.var_sub5$years_since)
env.var_sub5$years_since <- ifelse(env.var_sub5$year == "2011", "9", env.var_sub5$years_since)
env.var_sub5$years_since <- ifelse(env.var_sub5$year == "2012", "8", env.var_sub5$years_since)
env.var_sub5$years_since <- ifelse(env.var_sub5$year == "2014", "6", env.var_sub5$years_since)
env.var_sub5$years_since <- ifelse(env.var_sub5$year == "2016", "4", env.var_sub5$years_since)
str(env.var_sub5$years_since)
env.var_sub5$years_since <- as.numeric(env.var_sub5$years_since)

str(env.var_sub5)

save(env.var_sub5, file = "env.var_sub5_095_k4_2020_09_10.rda")


# To check if we get other results when keeping most of the observations (we exclude a lot of reference plots when doing species_richness >5)
levels(env.var_sub1$year)

env.var_sub1$years_since <- ifelse(env.var_sub1$year == "2003", "17", NA)
env.var_sub1$years_since <- ifelse(env.var_sub1$year == "2007", "13", env.var_sub1$years_since)
env.var_sub1$years_since <- ifelse(env.var_sub1$year == "2009", "11", env.var_sub1$years_since)
env.var_sub1$years_since <- ifelse(env.var_sub1$year == "2011", "9", env.var_sub1$years_since)
env.var_sub1$years_since <- ifelse(env.var_sub1$year == "2012", "8", env.var_sub1$years_since)
env.var_sub1$years_since <- ifelse(env.var_sub1$year == "2014", "6", env.var_sub1$years_since)
env.var_sub1$years_since <- ifelse(env.var_sub1$year == "2016", "4", env.var_sub1$years_since)
str(env.var_sub1$years_since)
env.var_sub1$years_since <- as.numeric(env.var_sub1$years_since)

str(env.var_sub1)

save(env.var_sub1, file = "env.var_sub1_095_k4_2020_09_11.rda")

```







## Make dataset with all explanatory variables of env.var_full_subs to use for the models
```{r}
str(env.var_full_sub5$grain_size_stand_f)
table(env.var_full_sub5$grain_size_stand_f)
 #  1   2   3   5   6
 # 27   0  13   2 177
env.var_full_sub5$grain_size_stand_f <- as.numeric(env.var_full_sub5$grain_size_stand_f)
env.var_full_sub5$grain_size_stand_f <- ifelse(env.var_full_sub5$grain_size_stand_f >= 4, 6, env.var_full_sub5$grain_size_stand_f)
table(env.var_full_sub5$grain_size_stand_f)
 #  1   3   6
 # 27  13 179
env.var_full_sub5$grain_size_stand_f <- as.factor(env.var_full_sub5$grain_size_stand_f)
str(env.var_full_sub5$grain_size_stand_f)
table(env.var_full_sub5$grain_size_stand_f)


str(env.var_full_sub5$exposure)
table(env.var_full_sub5$exposure)
  # E   F   N  NE  NW   S  SE  SW   W
  # 9 124  31   1   6  19  13   6  10
levels(env.var_full_sub5$exposure)[levels(env.var_full_sub5$exposure)=="NE"] <- "N"
str(env.var_full_sub5)
table(env.var_full_sub5$exposure)


save(env.var_full_sub5, file = "env.var_full_sub5_095_k4_2020_09_11.rda")
```


```{r}
str(env.var_full_sub1$grain_size_stand_f)
table(env.var_full_sub1$grain_size_stand_f)
 #  1   2   3   5   6 
 # 33   1  16   2 230 
env.var_full_sub1$grain_size_stand_f <- as.numeric(env.var_full_sub1$grain_size_stand_f)
env.var_full_sub1$grain_size_stand_f <- ifelse(env.var_full_sub1$grain_size_stand_f >= 4, 6, env.var_full_sub1$grain_size_stand_f)
table(env.var_full_sub1$grain_size_stand_f)
 #  1   2   3   6 
 # 33   1  16 232 
env.var_full_sub1$grain_size_stand_f <- ifelse(env.var_full_sub1$grain_size_stand_f == 2, 1, env.var_full_sub1$grain_size_stand_f)
env.var_full_sub1$grain_size_stand_f <- as.factor(env.var_full_sub1$grain_size_stand_f)
str(env.var_full_sub1$grain_size_stand_f)
table(env.var_full_sub1$grain_size_stand_f)
 #  1   3   4 
 # 34  16 232

str(env.var_full_sub1$exposure)
table(env.var_full_sub1$exposure)
 #  E   F   N  NE  NW   S  SE  SW   W 
 # 13 166  33   6   7  23  17   6  11 


save(env.var_full_sub1, file = "env.var_full_sub1_095_k4_2020_09_11.rda")
```


#### Change levels for method, so that reference is in intercept
```{r}
levels(env.var_full_sub1$method)
env.var_full_sub1$method <- factor(env.var_full_sub1$method, levels = c("ref", "nat",  "pn",  "seed"))
plot(env.var_full_sub1$method)

levels(env.var_full_sub1$method)
env.var_full_sub1$method <- factor(env.var_full_sub1$method, levels = c("ref", "nat",  "pn",  "seed"))
plot(env.var_full_sub1$method)

```


#### Make a variable years since restoration
```{r}
levels(env.var_full_sub5$year)

env.var_full_sub5$years_since <- ifelse(env.var_full_sub5$year == "2003", "17", NA)
env.var_full_sub5$years_since <- ifelse(env.var_full_sub5$year == "2007", "13", env.var_full_sub5$years_since)
env.var_full_sub5$years_since <- ifelse(env.var_full_sub5$year == "2009", "11", env.var_full_sub5$years_since)
env.var_full_sub5$years_since <- ifelse(env.var_full_sub5$year == "2011", "9", env.var_full_sub5$years_since)
env.var_full_sub5$years_since <- ifelse(env.var_full_sub5$year == "2012", "8", env.var_full_sub5$years_since)
env.var_full_sub5$years_since <- ifelse(env.var_full_sub5$year == "2014", "6", env.var_full_sub5$years_since)
env.var_full_sub5$years_since <- ifelse(env.var_full_sub5$year == "2016", "4", env.var_full_sub5$years_since)
str(env.var_full_sub5$years_since)
env.var_full_sub5$years_since <- as.numeric(env.var_full_sub5$years_since)

str(env.var_full_sub5)

save(env.var_full_sub5, file = "env.var_full_sub5_095_k4_2020_09_11.rda")



levels(env.var_full_sub1$year)

env.var_full_sub1$years_since <- ifelse(env.var_full_sub1$year == "2003", "17", NA)
env.var_full_sub1$years_since <- ifelse(env.var_full_sub1$year == "2007", "13", env.var_full_sub1$years_since)
env.var_full_sub1$years_since <- ifelse(env.var_full_sub1$year == "2009", "11", env.var_full_sub1$years_since)
env.var_full_sub1$years_since <- ifelse(env.var_full_sub1$year == "2011", "9", env.var_full_sub1$years_since)
env.var_full_sub1$years_since <- ifelse(env.var_full_sub1$year == "2012", "8", env.var_full_sub1$years_since)
env.var_full_sub1$years_since <- ifelse(env.var_full_sub1$year == "2014", "6", env.var_full_sub1$years_since)
env.var_full_sub1$years_since <- ifelse(env.var_full_sub1$year == "2016", "4", env.var_full_sub1$years_since)
str(env.var_full_sub1$years_since)
env.var_full_sub1$years_since <- as.numeric(env.var_full_sub1$years_since)

str(env.var_full_sub1)

save(env.var_full_sub1, file = "env.var_full_sub1_095_k4_2020_09_11.rda")

```




