---
title: "03_models"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ------------- head data ----------------
```{r}
head(data_plots2)
head(data_plots2_normal)
head(data_plots2_reference)

head(data_species_wide2)
head(data_species_wide2_normal)
head(data_species_wide2_reference)

```


## ----------- Test for data distribution/ screwness ---------
```{r}
library(fitdistrplus)

descdist(data_plots2_normal$field_layer_c, discrete = FALSE, boot = 500)
descdist(data_plots2_normal$shrub_layer_c, discrete = FALSE, boot = 500)
descdist(data_plots2_normal$tree_layer_c, discrete = FALSE, boot = 500)

descdist(data_plots3$total_cover_layers, discrete = FALSE, boot = 500)
descdist(data_plots3$mean_cover, discrete = FALSE, boot = 500)
descdist(data_plots3$total_cover_layers, discrete = TRUE, boot = 500)
descdist(data_plots3$mean_cover, discrete = TRUE, boot = 500)

descdist(data_plots2_reference$field_layer_c, discrete = FALSE, boot = 500)
descdist(data_plots2_reference$shrub_layer_c, discrete = FALSE, boot = 500)
descdist(data_plots2_reference$tree_layer_c, discrete = FALSE, boot = 500)

descdist(data_plots2_reference$total_cover_layers, discrete = FALSE, boot = 500)


descdist(data_plots3$field_layer_c, discrete = FALSE, boot = 500)
descdist(data_plots3$field_layer_c, discrete = TRUE, boot = 500)

```


## correlation
```{r}
cor_df <- data.frame(data_plots2_normal$bare_soil_c, data_plots2_normal$dist_int_veg, data_plots2_normal$field_layer_c, data_plots2_normal$shrub_layer_c, data_plots2_normal$tree_layer_c, data_plots2_normal$caco, data_plots2_normal$slope, data_plots2_normal$loi, data_plots2_normal$m, data_plots2_normal$sd_cover, data_plots2_normal$total_cover_layers)

res <- cor(cor_df)

library(Hmisc)
library(corrplot)
corrplot(res)

```



## -------------------- Linear mixed model with "total_cover_layers" as response (excl. reference plots) -------------------

```{r}
library(lme4)
library(lmerTest)

#### random factor selection ----
lm0_rand <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + site_type + (1|site_id/plot_id), data = data_plots2_normal, REML = F) # Error: number of levels of each grouping factor must be < number of observations (problems: plot_id:site_id)

lm1_rand <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method  + (1|site_id), data = data_plots2_normal, REML = F) #1

lm2_rand <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method  + (1|plot_id), data = data_plots2_normal, REML = F)



#### models ----
lm0 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + (1|site_id), data = data_plots2_normal, REML = F)
lm1 <- lmer(total_cover_layers ~ dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F)
lm2 <- lmer(total_cover_layers ~ dist_int_veg + caco + (1|site_id), data = data_plots2_normal, REML = F) 
lm3 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + (1|site_id), data = data_plots2_normal, REML = F)
lm4 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + (1|site_id), data = data_plots2_normal, REML = F)
lm5 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
lm6 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + (1|site_id), data = data_plots2_normal, REML = F)


lm7 <- lmer(total_cover_layers ~ dist_int_veg + method + (1|site_id), data = data_plots2_normal, REML = F)
lm8 <- lmer(total_cover_layers ~ dist_int_veg + method + caco + (1|site_id), data = data_plots2_normal, REML = F) 
lm9 <- lmer(total_cover_layers ~  method + (1|site_id), data = data_plots2_normal, REML = F) 

lm10 <- lmer(total_cover_layers ~ method*dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F) #2
lm11 <- lmer(total_cover_layers ~ method*dist_int_veg + caco + (1|site_id), data = data_plots2_normal, REML = F)
lm12 <- lmer(total_cover_layers ~ method*caco + (1|site_id), data = data_plots2_normal, REML = F) #1
lm12a_total_c <- lmer(total_cover_layers ~ method*caco + dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F)
lm13 <- lmer(total_cover_layers ~ method*caco + loi + (1|site_id), data = data_plots2_normal, REML = F) #3
lm14 <- lmer(total_cover_layers ~ method*caco + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
lm15 <- lmer(total_cover_layers ~ method*loi + (1|site_id), data = data_plots2_normal, REML = F)

```

#### random factor selection
```{r}
library(lmerTest)
AIC(lm1_rand, lm2_rand)
#          df      AIC
# lm1_rand 20 1970.204
# lm2_rand 20 1974.224
```



#### model selection 
```{r}
drop1(lm0)

s <- step(lm0)
plot(s)

library(AICcmodavg)
AIC(lm12, lm12a)

Cand.models <- list( )
Cand.models[[1]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[2]] <- lmer(total_cover_layers ~ dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[3]] <- lmer(total_cover_layers ~ dist_int_veg + caco + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[4]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[5]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[6]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[7]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[8]] <- lmer(total_cover_layers ~ dist_int_veg + method + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[9]] <- lmer(total_cover_layers ~ dist_int_veg + method + caco + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[10]] <- lmer(total_cover_layers ~ method + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[11]] <- lmer(total_cover_layers ~ method*dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[12]] <- lmer(total_cover_layers ~ method*dist_int_veg + caco + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[13]] <- lmer(total_cover_layers ~ method*caco + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[14]] <- lmer(total_cover_layers ~ method*caco + loi + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[15]] <- lmer(total_cover_layers ~ method*caco + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[16]] <- lmer(total_cover_layers ~ method*loi + (1|site_id), data = data_plots2_normal, REML = F)


aictab(Cand.models, second.ord = F) 
# Model selection based on AIC:
# 
#        K     AIC Delta_AIC AICWt Cum.Wt       LL
# Mod13  8 2686.25      0.00  0.32   0.32 -1335.12
# Mod11  8 2687.34      1.09  0.18   0.50 -1335.67
# Mod14  9 2688.21      1.96  0.12   0.62 -1335.10
# Mod16  8 2688.63      2.38  0.10   0.72 -1336.31
# Mod10  5 2688.70      2.45  0.09   0.81 -1339.35
# Mod12  9 2688.72      2.47  0.09   0.91 -1335.36
# Mod8   6 2689.80      3.55  0.05   0.96 -1338.90
# Mod9   7 2691.70      5.46  0.02   0.98 -1338.85
# Mod15 11 2691.95      5.71  0.02   1.00 -1334.98
# Mod2   4 2701.73     15.48  0.00   1.00 -1346.87
# Mod3   5 2703.14     16.89  0.00   1.00 -1346.57
# Mod4   6 2703.98     17.73  0.00   1.00 -1345.99
# Mod1  20 2707.36     21.12  0.00   1.00 -1333.68
# Mod5  14 2713.71     27.47  0.00   1.00 -1342.86
# Mod6  17 2719.65     33.40  0.00   1.00 -1342.82
# Mod7  18 2720.90     34.65  0.00   1.00 -1342.45

summary(lm0) #full

summary(lm13) #3
summary(lm10) #2
summary(lm12) #1
```

#### Checking assumptions
```{r}
library(sjPlot)
plot_model(lm13, type = "diag")
plot_model(lm10, type = "diag")
plot_model(lm12, type = "diag")
```










## -------------- Linear mixed model with "m" as response (excl. reference plots) ------------------

```{r}
library(lme4)
library(lmerTest)

#### random factor selection ----
lm3_rand <- lmer(m ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + site_type + (1|site_id/plot_id), data = data_plots2_normal, REML = F) # Error: number of levels of each grouping factor must be < number of observations (problems: plot_id:site_id)

lm4_rand <- lmer(m ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method  + (1|site_id), data = data_plots2_normal, REML = F) #1

lm5_rand <- lmer(m ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method  + (1|plot_id), data = data_plots2_normal, REML = F)



#### models ----
lm16 <- lmer(m ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + (1|site_id), data = data_plots2_normal, REML = F)
lm16a <- lmer(m ~ log(dist_int_veg+1) + caco + slope + exposure + grain_size_stand_f + log(loi+1) + method + (1|site_id), data = data_plots2_normal, REML = F)
lm17 <- lmer(m ~ method + (1|site_id), data = data_plots2_normal, REML = F) 
lm18 <- lmer(m ~ method + exposure + (1|site_id), data = data_plots2_normal, REML = F) 
lm19 <- lmer(m ~ method + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
lm20 <- lmer(m ~ method + loi + (1|site_id), data = data_plots2_normal, REML = F) 

lm21 <- lmer(m ~ method*loi + (1|site_id), data = data_plots2_normal, REML = F) #2
lm22 <- lmer(m ~ method*grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
lm23_mean_c <- lmer(m ~ method*log(loi+1) + (1|site_id), data = data_plots2_normal, REML = F) #1
lm24 <- lmer(m ~ method*log(loi+1) + log(dist_int_veg+1) + (1|site_id), data = data_plots2_normal, REML = F) #3

lm25 <- lmer(log(m+1) ~ method*log(loi+1) + (1|site_id), data = data_plots2_normal, REML = F)

lm21_REML <- lmer(m ~ method*log(loi+1) + (1|site_id), data = data_plots2_normal, REML = T) #1
lm25_REML <- lmer(log(m+1) ~ method*log(loi+1) + (1|site_id), data = data_plots2_normal, REML = T)

```

#### random factor selection
```{r}
library(lmerTest)
AIC(lm1_rand, lm2_rand)
#          df      AIC
# lm1_rand 20 1970.204
# lm2_rand 20 1974.224
```



#### model selection 
```{r}
drop1(lm16)

s <- step(lm16)
plot(s)

library(AICcmodavg)
# AIC(lm0, lm1, lm2, lm3, lm4, lm5, lm6)

Cand.models <- list( )
Cand.models[[1]] <- lmer(m ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[2]] <- lmer(m ~ method + (1|site_id), data = data_plots2_normal, REML = F) 
Cand.models[[3]] <- lmer(m ~ method + exposure + (1|site_id), data = data_plots2_normal, REML = F) 
Cand.models[[4]] <- lmer(m ~ method + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[5]] <- lmer(m ~ method + loi + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[6]] <- lmer(m ~ method*loi + (1|site_id), data = data_plots2_normal, REML = F) #2
Cand.models[[7]] <- lmer(m ~ method*grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[8]] <- lmer(m ~ method*log(loi+1) + (1|site_id), data = data_plots2_normal, REML = F)#1
Cand.models[[9]] <- lmer(m ~ log(dist_int_veg+1) + caco + slope + exposure + grain_size_stand_f + log(loi+1) + method + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[10]] <- lmer(m ~ method*log(loi+1) + log(dist_int_veg+1) + (1|site_id), data = data_plots2_normal, REML = F) #3

aictab(Cand.models, second.ord = F) 
#        K    AIC Delta_AIC AICWt Cum.Wt    LL
# Mod8   8 -31.63      0.00  0.47   0.47 23.81
# Mod6   8 -30.39      1.24  0.25   0.72 23.20
# Mod10  9 -30.07      1.56  0.21   0.93 24.03
# Mod5   6 -26.73      4.90  0.04   0.97 19.36
# Mod3  13 -24.45      7.18  0.01   0.98 25.22
# Mod2   5 -24.25      7.38  0.01   1.00 17.13
# Mod9  20 -20.11     11.52  0.00   1.00 30.05
# Mod1  20 -19.75     11.88  0.00   1.00 29.87
# Mod4   8 -19.20     12.42  0.00   1.00 17.60
# Mod7  13 -18.84     12.79  0.00   1.00 22.42


summary(lm16) #full

summary(lm21_REML)
anova(lm21_REML)

```

#### Checking assumptions
```{r}
library(sjPlot)
plot_model(lm21_REML, type = "diag")
plot_model(lm25, type = "diag")

```

#### Plot results
```{r}
plot_model(lm21_REML, type = "est", show.intercept = T)
plot_model(lm21_REML, type = "eff")

plot_model(lm25_REML, type = "est", show.intercept = T)
plot_model(lm25_REML, type = "eff")

```





########################## Not done yet!!!

## -------------------- Model with "field_layer_c" as response (all plots) -------------------

```{r}
library(lme4)
library(lmerTest)

#### random factor selection ----
lm0_rand <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + site_type + (1|site_id/plot_id), data = data_plots2_normal, REML = F) # Error: number of levels of each grouping factor must be < number of observations (problems: plot_id:site_id)

lm1_rand <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method  + (1|site_id), data = data_plots2_normal, REML = F) #1

lm2_rand <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method  + (1|plot_id), data = data_plots2_normal, REML = F)



#### models ----
lm0 <- glmmTMB(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + (1|site_id), data = data_plots3, family = beta_family())

lm0 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + (1|site_id), data = data_plots2_normal, REML = F)
lm1 <- lmer(total_cover_layers ~ dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F)
lm2 <- lmer(total_cover_layers ~ dist_int_veg + caco + (1|site_id), data = data_plots2_normal, REML = F) 
lm3 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + (1|site_id), data = data_plots2_normal, REML = F)
lm4 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + (1|site_id), data = data_plots2_normal, REML = F)
lm5 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
lm6 <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + (1|site_id), data = data_plots2_normal, REML = F)


lm7 <- lmer(total_cover_layers ~ dist_int_veg + method + (1|site_id), data = data_plots2_normal, REML = F)
lm8 <- lmer(total_cover_layers ~ dist_int_veg + method + caco + (1|site_id), data = data_plots2_normal, REML = F) 
lm9 <- lmer(total_cover_layers ~  method + (1|site_id), data = data_plots2_normal, REML = F) 

lm10 <- lmer(total_cover_layers ~ method*dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F) #2
lm11 <- lmer(total_cover_layers ~ method*dist_int_veg + caco + (1|site_id), data = data_plots2_normal, REML = F)
lm12 <- lmer(total_cover_layers ~ method*caco + (1|site_id), data = data_plots2_normal, REML = F) #1
lm12a <- lmer(total_cover_layers ~ method*caco + dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F)
lm13 <- lmer(total_cover_layers ~ method*caco + loi + (1|site_id), data = data_plots2_normal, REML = F) #3
lm14 <- lmer(total_cover_layers ~ method*caco + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
lm15 <- lmer(total_cover_layers ~ method*loi + (1|site_id), data = data_plots2_normal, REML = F)

```

#### random factor selection
```{r}
library(lmerTest)
AIC(lm1_rand, lm2_rand)
#          df      AIC
# lm1_rand 20 1970.204
# lm2_rand 20 1974.224
```



#### model selection 
```{r}
drop1(lm0)

s <- step(lm0)
plot(s)

library(AICcmodavg)
AIC(lm12, lm12a)

Cand.models <- list( )
Cand.models[[1]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[2]] <- lmer(total_cover_layers ~ dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[3]] <- lmer(total_cover_layers ~ dist_int_veg + caco + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[4]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[5]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[6]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[7]] <- lmer(total_cover_layers ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[8]] <- lmer(total_cover_layers ~ dist_int_veg + method + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[9]] <- lmer(total_cover_layers ~ dist_int_veg + method + caco + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[10]] <- lmer(total_cover_layers ~ method + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[11]] <- lmer(total_cover_layers ~ method*dist_int_veg + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[12]] <- lmer(total_cover_layers ~ method*dist_int_veg + caco + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[13]] <- lmer(total_cover_layers ~ method*caco + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[14]] <- lmer(total_cover_layers ~ method*caco + loi + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[15]] <- lmer(total_cover_layers ~ method*caco + grain_size_stand_f + (1|site_id), data = data_plots2_normal, REML = F)
Cand.models[[16]] <- lmer(total_cover_layers ~ method*loi + (1|site_id), data = data_plots2_normal, REML = F)


aictab(Cand.models, second.ord = F) 
# Model selection based on AIC:
# 
#        K     AIC Delta_AIC AICWt Cum.Wt       LL
# Mod13  8 2686.25      0.00  0.32   0.32 -1335.12
# Mod11  8 2687.34      1.09  0.18   0.50 -1335.67
# Mod14  9 2688.21      1.96  0.12   0.62 -1335.10
# Mod16  8 2688.63      2.38  0.10   0.72 -1336.31
# Mod10  5 2688.70      2.45  0.09   0.81 -1339.35
# Mod12  9 2688.72      2.47  0.09   0.91 -1335.36
# Mod8   6 2689.80      3.55  0.05   0.96 -1338.90
# Mod9   7 2691.70      5.46  0.02   0.98 -1338.85
# Mod15 11 2691.95      5.71  0.02   1.00 -1334.98
# Mod2   4 2701.73     15.48  0.00   1.00 -1346.87
# Mod3   5 2703.14     16.89  0.00   1.00 -1346.57
# Mod4   6 2703.98     17.73  0.00   1.00 -1345.99
# Mod1  20 2707.36     21.12  0.00   1.00 -1333.68
# Mod5  14 2713.71     27.47  0.00   1.00 -1342.86
# Mod6  17 2719.65     33.40  0.00   1.00 -1342.82
# Mod7  18 2720.90     34.65  0.00   1.00 -1342.45

summary(lm0) #full

summary(lm13) #3
summary(lm10) #2
summary(lm12) #1
```

#### Checking assumptions
```{r}
library(sjPlot)
plot_model(lm13, type = "diag")
plot_model(lm10, type = "diag")
plot_model(lm12, type = "diag")
```

