---
title: "05_species_richness_diversity"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(vegan)
library(MASS)
library(plyr)
library(tidyverse)

load("processdata/data_species_wide2_2020_07_15.rda")
load("processdata/data_species_wide2_sites_ref_2020_07_15.rda")
load("processdata/data_species_wide2_sites_nor_2020_07_15.rda")

head(data_species_wide2)
names(data_species_wide2)

#### Create dataframes grouped by sites and one for reference plots and one for normal plots
# data_species_wide2_sites_nor <- data_species_wide2 %>%
#   filter(reference == "no")
# 
# data_species_wide2_sites_ref <- data_species_wide2 %>%
#   filter(reference == "yes")



species_plots <- data.frame(data_species_wide2[, c(1, 4:192)])
str(species_plots)

species_sites <- data.frame(data_species_wide2[, c(2, 4:192)])
str(species_sites)

species_sites <- species_sites %>%
  group_by(site_id) %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, sum),
    )


species_sites_nor <- data.frame(data_species_wide2_sites_nor[, c(2, 4:192)])
str(species_sites_nor)

species_sites_nor <- species_sites_nor %>%
  group_by(site_id) %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, sum),
    )

species_sites_ref <- data.frame(data_species_wide2_sites_ref[, c(2, 4:192)])
str(species_sites_ref)

species_sites_ref <- species_sites_ref %>%
  group_by(site_id) %>%
  summarise(
    across(taraxacum_sp:lolium_perenne, sum),
    )

```

## Species richness (count of species)
```{r}
# sum up the number of non-zero entries per row (1)
# the first column is ignored [,-1] as it is a site name, not a species count.

sr_plot <- ddply(species_plots, ~ id, function(x) {
data.frame(richness = sum(x[-1]>0))
})


sr_site <- ddply(species_sites, ~ site_id, function(x) {
data.frame(richness = sum(x[-1]>0))
})

sr_site_nor <- ddply(species_sites_nor, ~ site_id, function(x) {
data.frame(richness = sum(x[-1]>0))
})

sr_site_ref <- ddply(species_sites_ref, ~ site_id, function(x) {
data.frame(richness = sum(x[-1]>0))
})

```


## Menhinick's index (to account for sampling effort)
...is simply the number of species (n) divided by the square-root of the total number of individuals (N)
```{r}
menhinick <- function(x) {sum(x>0)/sqrt(sum(x))
}

sr_plots_menhinick <- ddply(species_plots, ~id, function(x) {data.frame(sr_menhinick = menhinick(x[-1]))
  })


sr_sites_menhinick <- ddply(species_sites, ~site_id, function(x) {data.frame(sr_menhinick = menhinick(x[-1]))
  })

sr_sites_nor_menhinick <- ddply(species_sites_nor, ~site_id, function(x) {data.frame(sr_menhinick = menhinick(x[-1]))
  })

sr_sites_ref_menhinick <- ddply(species_sites_ref, ~site_id, function(x) {data.frame(sr_menhinick = menhinick(x[-1]))
  })

```


## Margalef's index (to account for sampling effort)
...is the number of species (n) minus 1 divided by the natural logarithm of the total number of individuals (N).
```{r}
margalef <- function(x) {(sum(x>0)-1)/log(sum(x))
  }
sr_plots_margalef <- ddply(species_plots, ~id, function(x) {data.frame(sr_margalef = margalef(x[-1]))
  })


sr_sites_margalef <- ddply(species_sites, ~site_id, function(x) {data.frame(sr_margalef = margalef(x[-1]))
  })

sr_sites_nor_margalef <- ddply(species_sites_nor, ~site_id, function(x) {data.frame(sr_margalef = margalef(x[-1]))
  })

sr_sites_ref_margalef <- ddply(species_sites_ref, ~site_id, function(x) {data.frame(sr_margalef = margalef(x[-1]))
  })

```


## Shannon-Wiener diversity Index
```{r}

diversity_plots_shannon <- ddply(species_plots, ~id,function(x) {data.frame(shannon = diversity(x[-1], index = "shannon"))
  })


diversity_sites_shannon <- ddply(species_sites, ~site_id,function(x) {data.frame(shannon = diversity(x[-1], index = "shannon"))
  })

diversity_sites_nor_shannon <- ddply(species_sites_nor, ~site_id,function(x) {data.frame(shannon = diversity(x[-1], index = "shannon"))
  })

diversity_sites_ref_shannon <- ddply(species_sites_ref, ~site_id,function(x) {data.frame(shannon = diversity(x[-1], index = "shannon"))
  })


```


## Simpson's diversity Index
```{r}

diversity_plots_simpsons <- ddply(species_plots, ~id,function(x) {data.frame(simpson = diversity(x[-1], index = "simpson"))
  })


diversity_sites_simpsons <- ddply(species_sites, ~site_id,function(x) {data.frame(simpson = diversity(x[-1], index = "simpson"))
  })

diversity_sites_nor_simpsons <- ddply(species_sites_nor, ~site_id,function(x) {data.frame(simpson = diversity(x[-1], index = "simpson"))
  })

diversity_sites_ref_simpsons <- ddply(species_sites_ref, ~site_id,function(x) {data.frame(simpson = diversity(x[-1], index = "simpson"))
  })


```


## Add SR and diversity to dataframes
```{r} 
data_species_wide3 <- data_species_wide2 %>% 
  left_join(diversity_plots_shannon) %>% 
  left_join(diversity_plots_simpsons) %>% 
  left_join(sr_plots_menhinick) %>%
  left_join(sr_plots_margalef) %>%
  as.data.frame()


data_species_wide3_sites <- data_species_wide2_sites %>% 
  left_join(diversity_sites_shannon) %>% 
  left_join(diversity_sites_simpsons) %>% 
  left_join(sr_sites_menhinick) %>%
  left_join(sr_sites_margalef) %>%
  as.data.frame()

data_species_wide3_sites_nor <- species_sites_nor %>% 
  left_join(diversity_sites_nor_shannon) %>% 
  left_join(diversity_sites_nor_simpsons) %>% 
  left_join(sr_sites_nor_menhinick) %>%
  left_join(sr_sites_nor_margalef) %>%
  as.data.frame()

data_species_wide3_sites_ref <- species_sites_ref %>% 
  left_join(diversity_sites_ref_shannon) %>% 
  left_join(diversity_sites_ref_simpsons) %>% 
  left_join(sr_sites_ref_menhinick) %>%
  left_join(sr_sites_ref_margalef) %>%
  as.data.frame()


```


## Save
```{r}
# Includes columns with mean and sd for species cover, species richness measures and species diversity measures (for both restored- and reference plots):
save(data_species_wide3, file = "data_species_wide3_2020_07_15.rda")


# Includes columns with mean and sd for species cover, species richness measures and species diversity measures; grouped by site id (for both normal- and reference plots):
save(data_species_wide3_sites, file = "data_species_wide3_sites_2020_07_15.rda")
# Includes columns with mean and sd for species cover; grouped by site id (for normal plots):
save(data_species_wide2_sites_nor, file = "data_species_wide2_sites_nor_2020_07_15.rda")
# Includes columns with mean and sd for species cover; grouped by site id (for reference plots):
save(data_species_wide2_sites_ref, file = "data_species_wide2_sites_ref_2020_07_15.rda")
# Includes columns with species richness measures and species diversity measures; grouped by site id (for normal plots):
save(data_species_wide3_sites_nor, file = "data_species_wide3_sites_nor_2020_07_16.rda")
# Includes columns with species richness measures and species diversity measures; grouped by site id (for reference plots):
save(data_species_wide3_sites_ref, file = "data_species_wide3_sites_ref_2020_07_16.rda")

```




