---
title: "05_species_richness_diversity"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(vegan)
library(MASS)
library(plyr)

load("processdata/data_species_wide2_normal_2020_06_29.rda")
head(data_species_wide2_normal)
names(data_species_wide2_normal)

species_plots <- data.frame(data_species_wide2_normal[, c(1, 4:192)])
str(species_plots)

species_sites <- data.frame(data_species_wide2_normal[, c(2, 4:192)])
str(species_sites)


```

## Species richness (count of species)
```{r}
# sum up the number of non-zero entries per row (1)
# the first column is ignored [,-1] as it is a site name, not a species count.

sr_plot <- ddply(species_plots, ~ id, function(x) {
data.frame(richness = sum(x[-1]>0))
})

sr_site <- ddply(species_sites, ~ site_id, function(x) {
data.frame(richness = sum(x[-1]>0))
})

```


## Menhinick's index (to account for sampling effort)
...is simply the number of species (n) divided by the square-root of the total number of individuals (N)
```{r}
menhinick <- function(x) {sum(x>0)/sqrt(sum(x))
}

sr_plots_menhinick <- ddply(species_plots, ~id, function(x) {data.frame(richness = menhinick(x[-1]))
  })

sr_sites_menhinick <- ddply(species_sites, ~site_id, function(x) {data.frame(richness = menhinick(x[-1]))
  })

```


## Margalef's index (to account for sampling effort)
...is the number of species (n) minus 1 divided by the natural logarithm of the total number of individuals (N).
```{r}
margalef <- function(x) {(sum(x>0)-1)/log(sum(x))
  }
sr_plots_margalef <- ddply(species_plots, ~id, function(x) {data.frame(richness = margalef(x[-1]))
  })

sr_sites_margalef <- ddply(species_sites, ~site_id, function(x) {data.frame(richness = margalef(x[-1]))
  })


```


## Shannon-Wiener diversity Index
```{r}

diversity_plots_shannon <- ddply(species_plots, ~id,function(x) {data.frame(shannon = diversity(x[-1], index = "shannon"))
  })

diversity_sites_shannon <- ddply(species_sites, ~site_id,function(x) {data.frame(shannon = diversity(x[-1], index = "shannon"))
  })
#### Need to summarize data per site in the species_sites dataframe, so that I get 29 sites and not plots...

```


## Simpson's diversity Index
```{r}

diversity_plots_simpsons <- ddply(species_plots, ~id,function(x) {data.frame(simpson = diversity(x[-1], index = "simpson"))
  })

diversity_sites_simpsons <- ddply(species_sites, ~site_id,function(x) {data.frame(simpson = diversity(x[-1], index = "simpson"))
  })
#### Need to summarize data per site in the species_sites dataframe, so that I get 29 sites and not plots...

```