---
title: "06_models_species_diversity"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## ------------- head data ----------------
```{r}

head(data_plots4)

```


<!-- ## ----------- Test for data distribution/ screwness --------- -->
<!-- ```{r} -->
<!-- library(fitdistrplus) -->

<!-- ## If discrete is TRUE, the represented distributions are the Poisson, negative binomial distributions, and the normal distribution to which previous discrete distributions may converge. If discrete is FALSE, these are uniform, normal, logistic, lognormal, beta and gamma distributions. -->

<!-- # discrete = FALSE -->
<!-- descdist(data_plots4$shannon, discrete = FALSE, boot = 500) # Normal, maybe lognormal -->
<!-- descdist(data_plots4$simpson, discrete = FALSE, boot = 500) # beta -->
<!-- descdist(data_plots4$sr_margalef, discrete = FALSE, boot = 500) # beta, maybe normal -->
<!-- descdist(data_plots4$sr_menhinick, discrete = FALSE, boot = 500) # lognormal -->

<!-- # discrete = TRUE -->
<!-- descdist(data_plots4$shannon, discrete = TRUE, boot = 500) # neg. binomial -->
<!-- descdist(data_plots4$simpson, discrete = TRUE, boot = 500) # neg. binomial -->
<!-- descdist(data_plots4$sr_margalef, discrete = TRUE, boot = 500) # normal or poisson -->
<!-- descdist(data_plots4$sr_menhinick, discrete = TRUE, boot = 500) # neg. binomial -->

<!-- ``` -->


## correlation
```{r}
# cor_df <- data.frame(data_plots4$bare_soil_c, data_plots4$dist_int_veg, data_plots4$field_layer_c, data_plots4$shrub_layer_c, data_plots4$tree_layer_c, data_plots4$caco, data_plots4$slope, data_plots4$loi, data_plots4$mean_cover, data_plots4$sd_cover, data_plots4$total_cover_layers, data_plots4$shannon, data_plots4$simpson, data_plots4$sr_margalef, data_plots4$sr_menhinick)
# 
# res <- cor(cor_df)
# 
# library(Hmisc)
# library(corrplot)
# corrplot(res)

```




<!-- ## -------------------- Linear mixed model with Ordi 1 gmnds axis scores from AXIS 1 as response (all plots) ------------------- -->
<!-- Axis scales retrieved from ordination with dataset species_sub5 epsilon 0.9 and k=3 (load("processdata/envvar_full_sub5_2020_08_31.rda")), then use column 43 and 44 for gmnds values -->
<!-- ```{r} -->
<!-- library(lme4) -->
<!-- library(lmerTest) -->
<!-- env.var_full_sub5 <- env.var_full_sub5[, c(1:44)] -->
<!-- str(env.var_full_sub5) -->

<!-- lmer_method1 <- lmer(mean_gnmds1 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_dist_int_veg1 <- lmer(mean_gnmds1 ~ dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_loi1 <- lmer(mean_gnmds1 ~ loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_caco1 <- lmer(mean_gnmds1 ~ caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_expo1 <- lmer(mean_gnmds1 ~ exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_grain_size1 <- lmer(mean_gnmds1 ~ grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_site_type1 <- lmer(mean_gnmds1 ~ site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_year1 <- lmer(mean_gnmds1 ~ year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->


<!-- AIC(lmer_method1, lmer_loi1, lmer_caco1, lmer_expo1, lmer_grain_size1, lmer_site_type1, lmer_year1) -->
<!-- anova(lmer_method, lmer_loi) -->
<!-- anova(lmer_method, lmer_caco) -->
<!-- anova(lmer_method, lmer_dist_int_veg) -->


<!-- lmerfull1 <- lmer(mean_gnmds1 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- summary(lmerfull) -->

<!-- m1_a1 <- lmer(mean_gnmds1 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- m2_a1 <- lmer(mean_gnmds1 ~ method + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- m3_a1 <- lmer(mean_gnmds1 ~ method + caco + loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # BEST -->
<!-- m4_a1 <- lmer(mean_gnmds1 ~ method + caco + loi + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- m5_a1 <- lmer(mean_gnmds1 ~ method + caco + loi + exposure + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- # interaction between year and method -->


<!-- AIC(lmerfull1, m1_a1, m2_a1, m3_a1, m4_a1, m5_a1) -->
<!-- #       df      AIC -->
<!-- # m1_a1  6 274.1463 -->
<!-- # m2_a1  7 251.9221 -->
<!-- # m3_a1  8 248.5197  # BEST -->
<!-- # m4_a1 16 255.9748 -->
<!-- # m5_a1 22 249.9681  # 2nd -->

<!-- library(AICcmodavg) -->

<!-- Cand.models <- list( ) -->
<!-- Cand.models[[1]] <- lmer(mean_gnmds1 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[2]] <- lmer(mean_gnmds1 ~ method + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[3]] <- lmer(mean_gnmds1 ~ method + caco + loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # BEST -->
<!-- Cand.models[[4]] <- lmer(mean_gnmds1 ~ method + caco + loi + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[5]] <- lmer(mean_gnmds1 ~ method + caco + loi + exposure + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->

<!-- aictab(Cand.models, second.ord = F)  -->



<!-- summary(m3_a1_reml) -->

<!-- m3_a1_reml <- lmer(mean_gnmds1 ~ method + caco + loi + (1|site_id_new.x), data = env.var_full_sub5, REML = T) -->

<!-- ``` -->

<!-- #### Checking assumptions -->
<!-- ```{r} -->
<!-- library(sjPlot) -->
<!-- plot_model(m3_a1_reml, type = "diag") -->

<!-- ``` -->


<!-- #### Plot results -->
<!-- ```{r} -->
<!-- plot_model(m3_a1_reml, type = "est", show.intercept = T) -->
<!-- plot_model(m3_a1_reml, type = "eff", terms = "method") -->
<!-- plot_model(m3_a1_reml, type = "eff", terms = "caco") -->
<!-- plot_model(m3_a1_reml, type = "eff", terms = "loi") -->

<!-- tab_model(m3_a1_reml, p.style = "stars") -->

<!-- ``` -->




<!-- ## -------------------- Linear mixed model with Ordi 1 gmnds axis scores from AXIS 2 as response (all plots) ------------------- -->
<!-- Axis scales retrieved from ordination with dataset species_sub5 epsilon 0.9 and k=3 -->
<!-- #### Model selection -->
<!-- ```{r} -->
<!-- lmer_method2 <- lmer(mean_gnmds2 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_dist_int_veg1 <- lmer(mean_gnmds2 ~ dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_loi2 <- lmer(mean_gnmds2 ~ loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_caco2 <- lmer(mean_gnmds2 ~ caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_expo2 <- lmer(mean_gnmds2 ~ exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_grain_size2 <- lmer(mean_gnmds2 ~ grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_site_type2 <- lmer(mean_gnmds2 ~ site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- lmer_year2 <- lmer(mean_gnmds2 ~ year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->


<!-- AIC(lmer_method2, lmer_loi2, lmer_caco2, lmer_expo2, lmer_grain_size2, lmer_site_type2, lmer_year2) -->
<!-- anova(lmer_caco2, lmer_loi2) -->
<!-- anova(lmer_site_type2, lmer_loi2) -->


<!-- lmerfull2 <- lmer(mean_gnmds2 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- summary(lmerfull) -->

<!-- m1_a2 <- lmer(mean_gnmds2 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # Best -->
<!-- m2_a2 <- lmer(mean_gnmds2 ~ method + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- m3_a2 <- lmer(mean_gnmds2 ~ method + caco + loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- m4_a2 <- lmer(mean_gnmds2 ~ method + caco + loi + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- m5_a2 <- lmer(mean_gnmds2 ~ method + caco + loi + exposure + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- m6_a2 <- lmer(mean_gnmds2 ~ method + caco + loi + site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- m7_a2 <- lmer(mean_gnmds2 ~ method + caco + loi + site_type + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- m8_a2 <- lmer(mean_gnmds2 ~ method + caco + loi + site_type + exposure + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->


<!-- AIC(m1_a2, m2_a2, m3_a2, m4_a2, m5_a2, m6_a2, m7_a2, m8_a2) -->
<!-- #       df      AIC -->
<!-- # m1_a2  6 322.3958  # BEST -->
<!-- # m2_a2  7 323.4603 -->
<!-- # m3_a2  8 323.3754  # 2nd -->
<!-- # m4_a2 16 330.8082 -->
<!-- # m5_a2 22 335.9778 -->
<!-- # m6_a2  9 325.3553 -->
<!-- # m7_a2 17 332.8054 -->
<!-- # m8_a2 23 337.9731 -->

<!-- library(AICcmodavg) -->

<!-- Cand.models <- list( ) -->
<!-- Cand.models[[1]] <- lmer(mean_gnmds2 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[2]] <- lmer(mean_gnmds2 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # Best -->
<!-- Cand.models[[3]] <- lmer(mean_gnmds2 ~ method + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[4]] <- lmer(mean_gnmds2 ~ method + caco + loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[5]] <- lmer(mean_gnmds2 ~ method + caco + loi + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[6]] <- lmer(mean_gnmds2 ~ method + caco + loi + exposure + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[7]] <- lmer(mean_gnmds2 ~ method + caco + loi + site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[8]] <- lmer(mean_gnmds2 ~ method + caco + loi + site_type + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->
<!-- Cand.models[[9]] <- lmer(mean_gnmds2 ~ method + caco + loi + site_type + exposure + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) -->

<!-- aictab(Cand.models, second.ord = F)  -->


<!-- summary(m1_a2) -->
<!-- summary(m3_a2) -->

<!-- m1_a2_reml <- lmer(mean_gnmds2 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = T)  -->

<!-- ``` -->

<!-- #### Checking assumptions -->
<!-- ```{r} -->
<!-- library(sjPlot) -->
<!-- plot_model(m1_a2_reml, type = "diag") -->

<!-- ``` -->


<!-- #### Plot results -->
<!-- ```{r} -->
<!-- plot_model(m1_a2_reml, type = "est", show.intercept = T) -->
<!-- plot_model(m1_a2_reml, type = "eff", terms = "method") -->

<!-- tab_model(m1_a2_reml, p.style = "stars") -->

<!-- ``` -->




## -------------------- Linear mixed model with Ordi 2 gmnds axis scores from AXIS 1 as response (all plots) -------------------
Axis scales retrieved from ordination with dataset species_sub5 epsilon 0.95 and k=4 (load("processdata/env.var_full_sub5_095_k4_2020_09_10.rda"))

```{r}
str(env.var_full_sub5)
env.var_full_sub5$site_id_new.x <- as.factor(env.var_full_sub5$site_id_new.x)

```

```{r}
library(lme4)
library(lmerTest)
lmer_method1 <- lmer(mean_gnmds1 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_dist_int_veg1 <- lmer(mean_gnmds1 ~ dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_loi1 <- lmer(mean_gnmds1 ~ loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_caco1 <- lmer(mean_gnmds1 ~ caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_expo1 <- lmer(mean_gnmds1 ~ exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_grain_size1 <- lmer(mean_gnmds1 ~ grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_site_type1 <- lmer(mean_gnmds1 ~ site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_year1 <- lmer(mean_gnmds1 ~ year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)


AIC(lmer_method1, lmer_dist_int_veg1, lmer_loi1, lmer_caco1, lmer_expo1, lmer_grain_size1, lmer_site_type1, lmer_year1)
anova(lmer_method, lmer_loi)
anova(lmer_method, lmer_caco)
anova(lmer_method, lmer_dist_int_veg)


lmerfull1 <- lmer(mean_gnmds1 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
summary(lmerfull1)

m1_a1 <- lmer(mean_gnmds1 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m2_a1 <- lmer(mean_gnmds1 ~ method + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m3_a1 <- lmer(mean_gnmds1 ~ method + caco + loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 
m4_a1 <- lmer(mean_gnmds1 ~ method + caco + loi + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m5_a1 <- lmer(mean_gnmds1 ~ method + caco + loi + exposure + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m6_a1 <- lmer(mean_gnmds1 ~ method + caco + loi + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # BEST
m7_a1 <- lmer(mean_gnmds1 ~ method*year + caco + loi +  (1|site_id_new.x), data = env.var_full_sub5, REML = F) # Best, but what's with the warning???
# fixed-effect model matrix is rank deficient so dropping 10 columns / coefficients
m8_a1 <- lmer(mean_gnmds1 ~ method + log(caco+1) + log(loi+1) + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m9_a1 <- lmer(mean_gnmds1 ~ method + caco + log(loi+1) + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m10_a1 <- lmer(mean_gnmds1 ~ method + caco + loi + year + dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F)



AIC(lmerfull1, m1_a1, m2_a1, m3_a1, m4_a1, m5_a1, m6_a1, m7_a1)
AIC(m6_a1, m8_a1, m9_a1, m10_a1)
#       df      AIC 
#lmerfull1 27 196.7060
#   m1_a1  6 219.8779
#   m2_a1  7 203.1961
#   m3_a1  8 200.6369
#   m4_a1 16 202.4572
#   m5_a1 22 191.9755
#   m6_a1 14 188.0477 # BEST
#   m7_a1 22 163.6679

#       df      AIC
# m6_a1 14 188.0477 # BEST 
# m8_a1 14 192.3133
# m9_a1 14 190.9339
# m10_a1 15 189.7969

library(AICcmodavg)

Cand.models <- list( )
Cand.models[[1]] <- lmer(mean_gnmds1 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[2]] <- lmer(mean_gnmds1 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[3]] <- lmer(mean_gnmds1 ~ method + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[4]] <- lmer(mean_gnmds1 ~ method + caco + loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 
Cand.models[[5]] <- lmer(mean_gnmds1 ~ method + caco + loi + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[6]] <- lmer(mean_gnmds1 ~ method + caco + loi + exposure + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[7]] <- lmer(mean_gnmds1 ~ method + caco + loi + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # BEST
Cand.models[[8]] <- lmer(mean_gnmds1 ~ method*year + caco + loi +  (1|site_id_new.x), data = env.var_full_sub5, REML = F) # Best, but what's with the warning???
# fixed-effect model matrix is rank deficient so dropping 10 columns / coefficients
Cand.models[[9]] <- lmer(mean_gnmds1 ~ method + log(caco+1) + log(loi+1) + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[10]] <- lmer(mean_gnmds1 ~ method + caco + log(loi+1) + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[11]] <- lmer(mean_gnmds1 ~ method + caco + loi + year + dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F)

aictab(Cand.models, second.ord = F) 
#        K    AIC Delta_AIC AICWt Cum.Wt      LL
# Mod8  22 163.67      0.00     1      1  -59.83
# Mod7  14 188.05     24.38     0      1  -80.02
# Mod11 15 189.80     26.13     0      1  -79.90
# Mod6  21 190.29     26.62     0      1  -74.14
# Mod10 14 190.93     27.27     0      1  -81.47
# Mod9  14 192.31     28.65     0      1  -82.16
# Mod1  25 193.69     30.02     0      1  -71.84
# Mod4   8 200.64     36.97     0      1  -92.32
# Mod5  15 200.86     37.19     0      1  -85.43
# Mod3   7 203.20     39.53     0      1  -94.60
# Mod2   6 219.88     56.21     0      1 -103.94


m6_a1_reml <- lmer(mean_gnmds1 ~ method + caco + loi + year + (1|site_id_new.x), data = env.var_full_sub5, REML = T)

summary(m6_a1_reml)

```

#### Checking assumptions
```{r}
library(sjPlot)
plot_model(m6_a1_reml, type = "diag")

```


#### Plot results
```{r}
tab_model(m6_a1_reml, p.style = "stars")

plot_model(m6_a1_reml, type = "est", show.intercept = T)

plot_model(m6_a1_reml, type = "eff", terms = "method")
plot_model(m6_a1_reml, type = "eff", terms = "caco")
plot_model(m6_a1_reml, type = "eff", terms = "loi")
plot_model(m6_a1_reml, type = "eff", terms = "year")

```




## -------------------- Linear mixed model with gmnds axis scores from AXIS 2 as response (all plots) -------------------
Axis scales retrieved from ordination with dataset species_sub5 epsilon 0.95 and k=4 (load("processdata/env.var_full_sub5_095_k4_2020_09_10.rda"))
```{r}
lmer_method2 <- lmer(mean_gnmds2 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_dist_int_veg2 <- lmer(mean_gnmds2 ~ dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_loi2 <- lmer(mean_gnmds2 ~ loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_caco2 <- lmer(mean_gnmds2 ~ caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_expo2 <- lmer(mean_gnmds2 ~ exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_grain_size2 <- lmer(mean_gnmds2 ~ grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_site_type2 <- lmer(mean_gnmds2 ~ site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_year2 <- lmer(mean_gnmds2 ~ year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)


AIC(lmer_method2, lmer_dist_int_veg2, lmer_loi2, lmer_caco2, lmer_expo2, lmer_grain_size2, lmer_site_type2, lmer_year2)
anova(lmer_caco2, lmer_loi2)
anova(lmer_site_type2, lmer_loi2)


lmerfull2 <- lmer(mean_gnmds2 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
summary(lmerfull2)

m1_a2 <- lmer(mean_gnmds2 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 
m2_a2 <- lmer(mean_gnmds2 ~ method + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m3_a2 <- lmer(mean_gnmds2 ~ method + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m4_a2 <- lmer(mean_gnmds2 ~ method + year + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m5_a2 <- lmer(mean_gnmds2 ~ method + dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m6_a2 <- lmer(mean_gnmds2 ~ method + dist_int_veg + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m7_a2 <- lmer(mean_gnmds2 ~ method + year + caco + dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 
m8_a2 <- lmer(mean_gnmds2 ~ method + year + caco + dist_int_veg + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # Best

AIC(m1_a2, m2_a2, m3_a2, m4_a2, m5_a2, m6_a2, m7_a2, m8_a2)
#       df      AIC
# m1_a2  6 300.1943
# m2_a2 12 298.3096
# m3_a2 14 295.7121
# m4_a2 20 296.7170
# m5_a2  7 300.7337
# m6_a2  8 297.9013
# m7_a2 14 292.9709
# m8_a2 22 291.5535  # BEST

library(AICcmodavg)

Cand.models <- list( )
Cand.models[[1]] <- lmer(mean_gnmds2 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[2]] <- lmer(mean_gnmds2 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 
Cand.models[[3]] <- lmer(mean_gnmds2 ~ method + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[4]] <- lmer(mean_gnmds2 ~ method + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[5]] <- lmer(mean_gnmds2 ~ method + year + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[6]] <- lmer(mean_gnmds2 ~ method + dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[7]] <- lmer(mean_gnmds2 ~ method + dist_int_veg + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[8]] <- lmer(mean_gnmds2 ~ method + year + caco + dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 
Cand.models[[9]] <- lmer(mean_gnmds2 ~ method + year + caco + dist_int_veg + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F) #BEST

aictab(Cand.models, second.ord = F) 
#       K    AIC Delta_AIC AICWt Cum.Wt      LL
# Mod9 21 291.07      0.00  0.45   0.45 -124.53
# Mod1 25 292.77      1.70  0.19   0.65 -121.39
# Mod8 14 292.97      1.90  0.18   0.82 -132.49
# Mod4 13 294.24      3.18  0.09   0.92 -134.12
# Mod5 19 295.53      4.47  0.05   0.96 -128.77
# Mod7  8 297.90      6.83  0.01   0.98 -140.95
# Mod3 12 298.31      7.24  0.01   0.99 -137.15
# Mod2  6 300.19      9.13  0.00   1.00 -144.10
# Mod6  7 300.73      9.67  0.00   1.00 -143.37


summary(m8_a2)

m8_a2_reml <- lmer(mean_gnmds2 ~ method + year + caco + dist_int_veg + exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = T)

```


#### Checking assumptions
```{r}
library(sjPlot)
plot_model(m8_a2_reml, type = "diag")

```


#### Plot results
```{r}
tab_model(m8_a2_reml, p.style = "stars")

plot_model(m8_a2_reml, type = "est", show.intercept = T)

plot_model(m8_a2_reml, type = "eff", terms = "method")
plot_model(m8_a2_reml, type = "eff", terms = "caco")
plot_model(m8_a2_reml, type = "eff", terms = "exposure")
plot_model(m8_a2_reml, type = "eff", terms = "year")
plot_model(m8_a2_reml, type = "eff", terms = "dist_int_veg")

```





## -------------------- Linear mixed model with gmnds axis scores from AXIS 3 as response (all plots) -------------------
Axis scales retrieved from ordination with dataset species_sub5 epsilon 0.95 and k=4 (load("processdata/env.var_full_sub5_095_k4_2020_09_10.rda"))
```{r}
lmer_method3 <- lmer(mean_gnmds3 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_dist_int_veg3 <- lmer(mean_gnmds3 ~ dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_loi3 <- lmer(mean_gnmds3 ~ loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_caco3 <- lmer(mean_gnmds3 ~ caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_expo3 <- lmer(mean_gnmds3 ~ exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_grain_size3 <- lmer(mean_gnmds3 ~ grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_site_type3 <- lmer(mean_gnmds3 ~ site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_year3 <- lmer(mean_gnmds3 ~ year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)


AIC(lmer_method3, lmer_dist_int_veg3, lmer_loi3, lmer_caco3, lmer_expo3, lmer_grain_size3, lmer_site_type3, lmer_year3)
#                    df      AIC
# lmer_method3        6 246.0282
# lmer_dist_int_veg3  4 258.0863
# lmer_loi3           4 257.7306
# lmer_caco3          4 248.1830
# lmer_expo3         10 255.0127
# lmer_grain_size3    5 251.7939
# lmer_site_type3     4 256.3200
# lmer_year3          9 255.4212


lmerfull3 <- lmer(mean_gnmds3 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
summary(lmerfull3)

m1_a3 <- lmer(mean_gnmds3 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 
m2_a3 <- lmer(mean_gnmds3 ~ method + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m3_a3 <- lmer(mean_gnmds3 ~ method + grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m4_a3 <- lmer(mean_gnmds3 ~ method + caco + grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # BEST


AIC(m1_a3, m2_a3, m3_a3, m4_a3)
#       df      AIC
# m1_a3  6 246.0282
# m2_a3  7 244.7670
# m3_a3  8 241.0700
# m4_a3  9 238.8042


library(AICcmodavg)

Cand.models <- list( )
Cand.models[[1]] <- lmer(mean_gnmds3 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[2]] <- lmer(mean_gnmds3 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 
Cand.models[[3]] <- lmer(mean_gnmds3 ~ method + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[4]] <- lmer(mean_gnmds3 ~ method + grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[5]] <- lmer(mean_gnmds3 ~ method + caco + grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # BEST

aictab(Cand.models, second.ord = F) 
#       K    AIC Delta_AIC AICWt Cum.Wt      LL
# Mod5  9 238.80      0.00  0.67   0.67 -110.40
# Mod4  8 241.07      2.27  0.21   0.88 -112.53
# Mod1 25 243.38      4.57  0.07   0.95  -96.69
# Mod3  7 244.77      5.96  0.03   0.98 -115.38
# Mod2  6 246.03      7.22  0.02   1.00 -117.01


summary(m4_a3)

m4_a3_reml <- lmer(mean_gnmds3 ~ method + caco + grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F)

```


#### Checking assumptions
```{r}
library(sjPlot)
plot_model(m4_a3_reml, type = "diag")

```


#### Plot results
```{r}
tab_model(m4_a3_reml, p.style = "stars")

# plot_model(m4_a3_reml, type = "est", show.intercept = T)

plot_model(m4_a3_reml, type = "eff", terms = "method")
plot_model(m4_a3_reml, type = "eff", terms = "caco")
plot_model(m4_a3_reml, type = "eff", terms = "grain_size_stand_f")
```





## -------------------- Linear mixed model with gmnds axis scores from AXIS 4 as response (all plots) -------------------
Axis scales retrieved from ordination with dataset species_sub5 epsilon 0.95 and k=4 (load("processdata/env.var_full_sub5_095_k4_2020_09_10.rda"))
```{r}
lmer_method4 <- lmer(mean_gnmds4 ~ method + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_dist_int_veg4 <- lmer(mean_gnmds4 ~ dist_int_veg + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_loi4 <- lmer(mean_gnmds4 ~ loi + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_caco4 <- lmer(mean_gnmds4 ~ caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_expo4 <- lmer(mean_gnmds4 ~ exposure + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_grain_size4 <- lmer(mean_gnmds4 ~ grain_size_stand_f + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_site_type4 <- lmer(mean_gnmds4 ~ site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
lmer_year4 <- lmer(mean_gnmds4 ~ year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)


AIC(lmer_method4, lmer_dist_int_veg4, lmer_loi4, lmer_caco4, lmer_expo4, lmer_grain_size4, lmer_site_type4, lmer_year4)
                   # df      AIC
# lmer_method4        6 282.7611
# lmer_dist_int_veg4  4 286.7370
# lmer_loi4           4 285.2943
# lmer_caco4          4 280.3254
# lmer_expo4         10 292.5625
# lmer_grain_size4    5 291.9811
# lmer_site_type4     4 281.9206
# lmer_year4          9 293.5415


lmerfull4 <- lmer(mean_gnmds4 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
summary(lmerfull4)

m1_a4 <- lmer(mean_gnmds4 ~ site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 
m2_a4 <- lmer(mean_gnmds4 ~ method + site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m3_a4 <- lmer(mean_gnmds4 ~ caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
m4_a4 <- lmer(mean_gnmds4 ~ method + site_type + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # BEST
m5_a4 <- lmer(mean_gnmds4 ~ caco + site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 


AIC(m1_a4, m2_a4, m3_a4, m4_a4, m5_a4)
#       df      AIC
# m1_a4  4 281.9206
# m2_a4  7 282.4146
# m3_a4  4 280.3254
# m4_a4  8 273.3637
# m5_a4  5 274.7800


library(AICcmodavg)

Cand.models <- list( )
Cand.models[[1]] <- lmer(mean_gnmds4 ~ dist_int_veg + loi + caco + exposure + grain_size_stand_f + method + site_type + year + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[2]] <- lmer(mean_gnmds4 ~ method + site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[3]] <- lmer(mean_gnmds4 ~ caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
Cand.models[[4]] <- lmer(mean_gnmds4 ~ method + site_type + caco + (1|site_id_new.x), data = env.var_full_sub5, REML = F) # BEST
Cand.models[[5]] <- lmer(mean_gnmds4 ~ caco + site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F) 

aictab(Cand.models, second.ord = F) 
#       K    AIC Delta_AIC AICWt Cum.Wt      LL
# Mod4  8 273.36      0.00  0.65   0.65 -128.68
# Mod5  5 274.78      1.42  0.32   0.97 -132.39
# Mod3  4 280.33      6.96  0.02   0.99 -136.16
# Mod2  7 282.41      9.05  0.01   1.00 -134.21
# Mod1 25 293.78     20.42  0.00   1.00 -121.89


summary(m4_a4)

m4_a4_reml <- lmer(mean_gnmds4 ~ caco + site_type + (1|site_id_new.x), data = env.var_full_sub5, REML = F)
summary(m4_a4_reml)

```


#### Checking assumptions
```{r}
library(sjPlot)
plot_model(m4_a4_reml, type = "diag")

```


#### Plot results
```{r}
tab_model(m4_a4_reml, p.style = "stars")

# plot_model(m4_a4_reml, type = "est", show.intercept = T)

plot_model(m4_a4_reml, type = "eff", terms = "site_type")
plot_model(m4_a4_reml, type = "eff", terms = "caco")

```














## -------------------- Linear mixed model with "shannon" as response (all plots) -------------------

```{r}
library(lme4)
library(lmerTest)

plots_full <- data_full[, c(1, 3, 6:43, 47, 252)]
str(plots_full)
plots_full$method <- as.factor(plots_full$method)
plots_full$grain_size_stand_f <- as.factor(plots_full$grain_size_stand_f)
plots_full$exposure <- as.factor(plots_full$exposure)
plots_full$year <- as.factor(plots_full$year)


str(plots_full$grain_size_stand_f)
table(plots_full$grain_size_stand_f)
 #  1   2   3   5   6 
 # 33   1  16   2 233 
plots_full$grain_size_stand_f <- as.numeric(plots_full$grain_size_stand_f)
plots_full$grain_size_stand_f <- ifelse(plots_full$grain_size_stand_f >= 4, 6, plots_full$grain_size_stand_f)
plots_full$grain_size_stand_f <- ifelse(plots_full$grain_size_stand_f <= 2, 1, plots_full$grain_size_stand_f)
table(plots_full$grain_size_stand_f)
 #  1   3   6 
 # 34  16 235
plots_full$grain_size_stand_f <- as.factor(plots_full$grain_size_stand_f)
str(plots_full$grain_size_stand_f)
table(plots_full$grain_size_stand_f)


str(plots_full$exposure)
table(plots_full$exposure)
 #  E   F   N  NE  NW   S  SE  SW   W 
 # 14 168  33   6   7  23  17   6  11 


save(plots_full, file = "plots_full_shannon_2020_09_08.rda")




lmfull_sh <- lmer(shannon.x ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + year + region + (1|site_id_new.x), data = plots_full, REML = F) #fixed-effect model matrix is rank deficient so dropping 1 column / coefficient boundary (singular) fit: see ?isSingular
summary(lmfull_sh)

lm1_sh <- lmer(shannon.x ~ method + caco + exposure + grain_size_stand_f + (1|site_id_new.x), data = plots_full, REML = F) # BEST
lm2_sh <- lmer(shannon.x ~ method + caco + exposure + grain_size_stand_f + loi + (1|site_id_new.x), data = plots_full, REML = F) 
lm3_sh <- lmer(shannon.x ~ method + caco + exposure + grain_size_stand_f + year + (1|site_id_new.x), data = plots_full, REML = F) # boundary (singular) fit: see ?isSingular
lm4_sh <- lmer(shannon.x ~ method + caco + exposure + grain_size_stand_f + year + loi + (1|site_id_new.x), data = plots_full, REML = F) # boundary (singular) fit: see ?isSingular


lm1_sh_reml <- lmer(shannon.x ~  method + caco + exposure + grain_size_stand_f + (1|site_id_new.x), data = plots_full, REML = T)


```


#### model selection 
```{r}

AIC(lmfull_sh, lm1_sh, lm2_sh, lm3_sh)

library(AICcmodavg)

Cand.models <- list( )
Cand.models[[1]] <- lmer(shannon.x ~ dist_int_veg + caco + slope + exposure + grain_size_stand_f + loi + method + year + region + (1|site_id_new.x), data = plots_full, REML = F)
Cand.models[[2]] <- lmer(shannon.x ~ method + caco + exposure + grain_size_stand_f + (1|site_id_new.x), data = plots_full, REML = F) # BEST
Cand.models[[3]] <- lmer(shannon.x ~ method + caco + exposure + grain_size_stand_f + loi + (1|site_id_new.x), data = plots_full, REML = F) 
Cand.models[[4]] <- lmer(shannon.x ~ method + caco + exposure + grain_size_stand_f + year + (1|site_id_new.x), data = plots_full, REML = F) # boundary (singular) fit: see ?isSingular
Cand.models[[5]] <- lmer(shannon.x ~ method + caco + exposure + grain_size_stand_f + year + loi + (1|site_id_new.x), data = plots_full, REML = F)

aictab(Cand.models, second.ord = F) 
# Model selection based on AIC:
# 
#       K    AIC Delta_AIC AICWt Cum.Wt      LL
# Mod2 17 280.05      0.00  0.47   0.47 -123.02
# Mod3 18 280.16      0.11  0.44   0.91 -122.08
# Mod5 24 284.05      4.00  0.06   0.97 -118.02
# Mod4 23 286.61      6.56  0.02   0.99 -120.30
# Mod1 27 288.00      7.95  0.01   1.00 -117.00

summary(lm1_sh)
```

#### Checking assumptions
```{r}
library(sjPlot)
plot_model(lm1_sh_reml, type = "diag") ## Violates assumtion of normality of residuals... Do mixed models need that?


```


#### Plot results
```{r}
tab_model(lm1_sh_reml, p.style = "stars")

plot_model(lm1_sh_reml, type = "est", show.intercept = T)

plot_model(lm1_sh_reml, type = "eff", terms = "method")
plot_model(lm1_sh_reml, type = "eff", terms = "caco")
plot_model(lm1_sh_reml, type = "eff", terms = "exposure")
plot_model(lm1_sh_reml, type = "eff", terms = "grain_size_stand_f")


```







## -------------------- Linear mixed model with "sr_menhinick" as response (all plots) -------------------

```{r}
library(lme4)
library(lmerTest)

#### random factor selection ----
lm0_rand <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + field_layer_h + caco + slope + exposure + grain_size_stand_f + loi + method + site_type + (1|site_id/plot_id), data = data_plots4, REML = F) # Error: number of levels of each grouping factor must be < number of observations (problems: plotnumb:siteID)

lm1_rand <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + field_layer_h + caco + slope + exposure + grain_size_stand_f + loi + method + site_type + (1|site_id), data = data_plots4, REML = F) #1

lm2_rand <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + field_layer_h + caco + slope + exposure + grain_size_stand_f + loi + method + site_type +(1|plot_id), data = data_plots4, REML = F)



#### models ----
lm0 <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + field_layer_h + caco + slope + exposure + grain_size_stand_f + loi + method + (1|site_id), data = data_plots4, REML = F)

lm1 <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + caco + exposure + grain_size_stand_f + (1|site_id), data = data_plots4, REML = F)
lm2 <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + caco + grain_size_stand_f + (1|site_id), data = data_plots4, REML = F) 
lm3 <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + caco + grain_size_stand_f + log(loi+1) + (1|site_id), data = data_plots4, REML = F)
lm4_sr_menhinick <- lmer(sr_menhinick ~ log(dist_int_veg+1) + total_cover_layers + caco + grain_size_stand_f + log(loi+1) + (1|site_id), data = data_plots4, REML = F) #1
lm5 <- lmer(sr_menhinick ~ log(dist_int_veg+1) + total_cover_layers + caco + grain_size_stand_f + log(loi+1) + method + (1|site_id), data = data_plots4, REML = F) #3
lm6 <- lmer(sr_menhinick ~ log(dist_int_veg+1) + total_cover_layers + grain_size_stand_f + log(loi+1) + (1|site_id), data = data_plots4, REML = F) #2


lm4_reml <- lmer(sr_menhinick ~ log(dist_int_veg+1) + total_cover_layers  + caco + grain_size_stand_f + log(loi+1) + (1|site_id), data = data_plots4, REML = T) #1

```



#### random factor selection
```{r}
library(lmerTest)
AIC(lm1_rand, lm2_rand)
#          df      AIC
# lm1_rand 23 275.2039
# lm2_rand 23 282.9316

```



#### model selection 
```{r}
drop1(lm0)

s <- step(lm0)
plot(s)

library(AICcmodavg)
AIC(lm2, lm4)

Cand.models <- list( )
Cand.models[[1]] <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + field_layer_h + caco + slope + exposure + grain_size_stand_f + loi + method + (1|site_id), data = data_plots4, REML = F)
Cand.models[[2]] <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + caco + exposure + grain_size_stand_f + (1|site_id), data = data_plots4, REML = F)
Cand.models[[3]] <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers  + caco + grain_size_stand_f + (1|site_id), data = data_plots4, REML = F)
Cand.models[[4]] <- lmer(sr_menhinick ~ dist_int_veg + total_cover_layers + caco + grain_size_stand_f + log(loi+1) + (1|site_id), data = data_plots4, REML = F)
Cand.models[[5]] <- lmer(sr_menhinick ~ log(dist_int_veg+1) + total_cover_layers  + caco + grain_size_stand_f + log(loi+1) + (1|site_id), data = data_plots4, REML = F)
Cand.models[[6]] <- lmer(sr_menhinick ~ log(dist_int_veg+1) + total_cover_layers  + caco + grain_size_stand_f + log(loi+1) + method + (1|site_id), data = data_plots4, REML = F)
Cand.models[[7]] <- lmer(sr_menhinick ~ log(dist_int_veg+1) + total_cover_layers + grain_size_stand_f + log(loi+1) + (1|site_id), data = data_plots4, REML = F)


aictab(Cand.models, second.ord = F) 
# Model selection based on AIC:
# 
#       K    AIC Delta_AIC AICWt Cum.Wt     LL
# Mod5 11 125.05      0.00  0.49   0.49 -51.52
# Mod7 10 125.27      0.22  0.43   0.92 -52.64
# Mod6 13 128.98      3.94  0.07   0.99 -51.49
# Mod4 11 132.85      7.81  0.01   1.00 -55.43
# Mod3 10 136.47     11.42  0.00   1.00 -58.23
# Mod2 18 146.60     21.55  0.00   1.00 -55.30
# Mod1 23 151.84     26.79  0.00   1.00 -52.92


summary(lm0) #full

summary(lm6) #2
summary(lm4) #1
summary(lm4_reml) #1

```

#### Checking assumptions
```{r}
library(sjPlot)
plot_model(lm4_reml, type = "diag")

```


#### Plot results
```{r}
plot_model(lm4_reml, type = "est", show.intercept = T)
plot_model(lm4_reml, type = "eff")


```







